### 2023-07-09 paper/paper.Rmd
add files for writing paper

---
title: "testy"
output: pdf_document
date: "2023-07-09"
---

```{r dependencies, include=FALSE}
library(data.table)
library(here)
renv::snapshot()
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r read-data-all}
data <- fread(here("data", "clean", "all.csv"))
```

## Including Plots

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

\begin{equation}
\lim_{x\to 0} \frac{\exp(x)-1}{x} = 1
\end{equation}

Inline code example: `r data[, uniqueN(participant.code)]`

### 2023-07-09 paper/compile_paper.R

@@ -1,0 +1,8 @@
+# install missing packages from the lockfile
+renv::restore()
+
+# paper options
+knitr::opts_chunk$set(eval=T, echo=F, warning=F, message=F, cache=F)
+
+# compile the paper
+rmarkdown::render("paper/paper.Rmd")

### 2023-07-09 analysis/effects_aggregate.py
fixed file paths

@@ -22,7 +22,7 @@ plt.title('Performance by number of revealed variables')
 plt.axhline(.5, 0, 1, color = 'grey')
 axs.set_xlabel('number of revealed variables')
 axs.set_ylabel('share of correct guesses')
-fig.show()
+
 
 fig.savefig('computed_objects/figures/revealed_variables_pooled.png')
 
@@ -34,8 +34,6 @@ plt.title('Performance by number of available variables')
 plt.axhline(.5, 0, 1, color = 'grey')
 axs.set_xlabel('number of available variables')
 axs.set_ylabel('share of correct guesses')
-
-fig.show()
 
 fig.savefig('computed_objects/figures/assigned_variables.png')
 

### 2023-07-09 analysis/model_choices.py

@@ -74,4 +74,4 @@ plt.axhline(1, 0, 1, color = 'grey', linestyle='--', alpha = .5)
 plt.axhline(.85, 0, 1, color = 'grey', linestyle='--', alpha = .5)
 
 
-plt.savefig('computed_objects/figures/acuracy_rounds.png')
+plt.savefig('computed_objects/figures/accuracy_rounds.png')

### 2023-07-09 analysis/model_exploration.py

@@ -28,8 +28,7 @@ count_models.reset_index(inplace=True)
 
 counted_models = count_models.groupby(['player.number_variables', 'model']).count()
 counted_models.reset_index(inplace=True)
-print('the average number of models used by number of available variables is:')
-print(counted_models.groupby('player.number_variables')['model'].mean())
+
 
 count_models.rename(columns={'model':'number_models'}, inplace=True)
 # plot the histogram of the number of models used by number of available variables

### 2023-07-09 analysis/part2_performance.py

@@ -65,12 +65,6 @@ N = len(all_rounds['participant.code'].unique())
 
 share_better_random = better_random_count/N
 
-print('the number of subjects who do better than random in rounds 20 - 40 is ' + str(better_random_count))
-
-print('the number of subjects who do random or worse in rounds 20 - 40  is ' + str(N - better_random_count))
-
-print('the share of subjects who do better than random in rounds 20 - 40 is ' + str(share_better_random))
-
 
 # Accuracy wrt choosen model conditional on the number of variables they were allowed to choose by group in part 2 for 
 # one variable models and two variable models
@@ -98,8 +92,6 @@ mod = smf.ols(formula='y ~ indic_1 + indic_2+ indic_3 + indic_4 + indic_5  - 1',
 
 res_assigned = mod.fit(cov_type='HC3')
 
-print(res_assigned.summary())
-
 stargazer = Stargazer([res_assigned], )
 
 stargazer.render_latex()
@@ -108,7 +100,6 @@ stargazer.render_latex()
 # test that 1 variable makes the guesses better than random, and whether adding an extra variable is helpful or not  
 hypotheses = 'indic_1-.5 = 0, indic_2 - indic_1 = 0, indic_3 - indic_2 = 0, indic_4 - indic_3 = 0, indic_5 - indic_4 = 0'
 t_test = res_assigned.t_test(hypotheses)
-print(t_test)
 
 fig, axs = plt.subplots(2, 1, figsize=(25, 15))
 sns.pointplot(data=part2[(part2['player.number_variables']==2) ], x='model', y='player.correct',
@@ -133,4 +124,4 @@ part2 = part2.merge(model_accuracy, on='model', how='outer')
 sns.lineplot(data=part2[(part2['player.number_variables']==2) ], x='model', y='accuracy', ax=axs[0], 
               color='black')
 
-fig.savefig('computed_objects/figures/p2_perfotmance_model_choices_2var.png')
+fig.savefig('computed_objects/figures/p2_performance_model_choices_2var.png')

### 2023-07-09 analysis/polarization.py

@@ -134,8 +134,6 @@ mod = smf.ols(formula='y ~ predicted_polarization+round_number_modif+round_numbe
 
 res_predicted = mod.fit(cov_type='HC3')
 
-print(res_predicted.summary())
-
 stargazer = Stargazer([res_predicted], )
 
 with open('computed_objects/tables/regression_predicted_polariz.txt', 'w') as f:
@@ -143,7 +141,6 @@ with open('computed_objects/tables/regression_predicted_polariz.txt', 'w') as f:
 
 hypotheses = 'round_number_modif-round_number_modif:predicted_polarization=0'
 t_test = res_predicted.t_test(hypotheses)
-print(t_test.summary())
 
 with open('computed_objects/tables/ttest_predicted_polariz.txt', 'w') as f:
     f.write(str(t_test.summary()))
@@ -180,7 +177,6 @@ mod = smf.ols(formula='y ~1+ C(predicted_polarization)', data=pairs)
 
 res_polariz = mod.fit(cov_type='HC3')
 
-print(res_polariz.summary())
 
 stargazer = Stargazer([res_polariz], )
 

### 2023-07-09 analysis/regressions.py

@@ -26,8 +26,6 @@ mod = smf.ols(formula='y ~ indic_1 + indic_2+ indic_3 + indic_4 + indic_5  - 1',
 
 res_assigned = mod.fit(cov_type='HC3')
 
-print(res_assigned.summary())
-
 stargazer = Stargazer([res_assigned], )
 
 with open('computed_objects/tables/regression_assigned.txt', 'w') as f:
@@ -39,7 +37,7 @@ with open('computed_objects/tables/regression_assigned.txt', 'w') as f:
 # test that 1 variable makes the guesses better than random, and whether adding an extra variable is helpful or not  
 hypotheses = 'indic_1-.5 = 0, indic_2 - indic_1 = 0, indic_3 - indic_2 = 0, indic_4 - indic_3 = 0, indic_5 - indic_4 = 0'
 t_test_assigned = res_assigned.t_test(hypotheses)
-print(t_test_assigned)
+
 with open('computed_objects/tables/ttest_assigned.txt', 'w') as f:
     f.write(str(t_test_assigned.summary()))
 
@@ -51,7 +49,6 @@ mod = smf.ols(formula='y ~ C(revealed_variables_count)-1', data=part2)
 
 res_revealed = mod.fit(cov_type='HC3')
 
-print(res_revealed.summary())
 
 stargazer = Stargazer([res_revealed], )
 
@@ -61,7 +58,7 @@ with open('computed_objects/tables/regression_revealed.txt', 'w') as f:
 
 hypotheses = 'C(revealed_variables_count)[0] - .5 = 0, C(revealed_variables_count)[1] - C(revealed_variables_count)[0] = 0, C(revealed_variables_count)[2] - C(revealed_variables_count)[1] = 0, C(revealed_variables_count)[3] - C(revealed_variables_count)[2] = 0, C(revealed_variables_count)[4] - C(revealed_variables_count)[3] = 0, C(revealed_variables_count)[5] - C(revealed_variables_count)[4] = 0'
 t_test_revealed = res_revealed.t_test(hypotheses)
-print(t_test_revealed)
+
 
 with open('computed_objects/tables/ttest_revealed.txt', 'w') as f:
     f.write(str(t_test_revealed.summary()))

### 2023-07-09 analysis/variable_choices.py

@@ -72,6 +72,4 @@ axs[2].set_ylim(0. ,.8)
 axs[2].set_xlabel('variable')
 axs[2].set_xticks([0, 1, 2, 3, 4], ['x1', 'x2', 'x3', 'x4', 'x5'])
 
-fig.show()
-
 plt.savefig('computed_objects/figures/variable_choices_time.png')

### 2023-07-11 paper/paper.Rmd
beginning of experimental proc.

@@
 ---
-title: "testy"
-output: pdf_document
-date: "2023-07-09"
+title: "Learning from Data Through Models"
+date: '`r format(Sys.time(), "%B %d, %Y")`'
+author: "Alberto Bisin, Guillaume Frechette and Jimena Galindo"
+abstract: "TBW"
+output: 
+    pdf_document:
+        citation_package: natbib
+fontsize: 12pt
+geometry:
+- margin=2.5cm
+hyperref:
+- linestretch=2.0
+documentclass: article
 ---
@@
 renv::snapshot()
+
 ```
 
-## R Markdown
+# Introduction
+TBW
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+# Literature Review
+TBW
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+# Framework and Predictions
 
+A finite number of observable random variables, $X =(x_1, x_2, ..., x_N)$, determine a binary state, $y$.
+The observable variables are independently distributed and each follows a Normal distribution with corresponding mean $\mu_i$ and standard deviation $\sigma_i$ with $i \in \{1, 2, ..., N\}$.
+
+The state is \emph{Red} (**R**) or \emph{Blue} (**B**) and is determined in the following way:
+\begin{equation*}
+y = \begin{cases}
+\text{\textbf{B}} &\text{if $a_1x_1+...+a_Nx_N + K\geq0$}\\
+\text{\textbf{R}} &\text{otherwise}
+\end{cases}
+\end{equation*}
+
+An observable variable $x_i$ is *relevant* if the associated coefficient, $a_i$, is non-zero. 
+Likewise, variable $i$ is *irrelevant* if $a_i = 0$. 
+We call the set of relevant variables $\mathcal{R}$ and the set of irrelevant variables $\mathcal{I}$.
+
+In each period, the decision-maker (DM) observes the realization of all observable variables and predicts the state. 
+Their flow payoff in period $t$ is 1 if they predict the state correctly, and it is zero otherwise. 
+Denote the prediction by $\hat{y}$.
+
+In order to make a prediction, the DM uses a map $f: M_t\to \{\textbf{B}, \textbf{R}\}$  
+where $M_t \subseteq \{x_1, x_2, ..., x_N\}$ is the set of variables that they choose to consider in period $t$. 
+Our main object of study is the set of variables that the DM decides to consider and not the properties of the 
+mapping from the variables to the state. We refer to the set of variables that are taken into consideration, $M_t$, 
+as the DM's *model of the world*, and we abstract away from how the variables are mapped into a prediction of 
+the state.
+
+A model of the world is said to be *simple* if it does not include all the relevant variables, 
+*i.e.* if $M\subset \mathcal{R}$. And it is complex if it includes all relevant variables, $M=\mathcal{R}$.
+
+## Optimal Rules
+The most accurate prediction rule for this setting will be one that predicts the state to be **B** whenever, 
+given the observables, **B** is more likely. And it predicts **R** whenever **R** is more likely 
+given the observables. Because we abstract away from the procedure through which the rule is updated, 
+we will make predictions under the assumption that the DM has learned the parameters perfectly. That is, 
+they have seen enough information that, under an updating procedure that converges to the truth---for example, 
+correctly specified regression model or Bayes rule---they would have converged to the true parameters already.
+
+We refer to a prediction rule as an *Optimal Rule* if it is the rule that a fully rational Decision maker would 
+converge to with perfect Bayesian learning. In this context, Bayesian learning corresponds to using Bayes rule to 
+learn the coefficients $a = (a_1, a_2, ..., a_N)$ as well as the constant $K$. This means that in an $N$ dimensional 
+space, the Decision Maker must learn $N+1$ parameters which corresponds to learning the halfspace in which the state 
+is **B**. The optimal rule for the complex model coincides with the true data-generating process when learning 
+is perfect.
+
+Conditional on using a simple model, the optimal rule for that model is the rule that, when restricted to the 
+variables being considered, does the best at predicting the state. In this case, the Decision Maker is learning only 
+the parameters that pertain the variables in the model. Hence, for a model that considers $K$ variables, the DM will 
+have to learn $K+1$ parameters. The following proposition illustrates exactly what the optimal prediction rules look 
+like when the DM is able to learn.
+
+##### Proposition.
+_Fix any model that considers $k\leq N$ variables and relabel the variable so that $x_1, ..., x_K$ are the variables 
+that the model considers and $x_{k+1}, ..., x_N$ are the ones that the model does not consider. Relabel the 
+coefficients $a = (a_1, ..., a_N)$ accordingly as well._
+
+##### proof.
+_Let $M:=a_1x_1+...+a_kx_k$ and $M^c = a_{k+1}x_{k+1}+...+a_Nx_N+k$ and define the latent variable $y^L := M +  M^c$. 
+The optimal prediction rule for a model that considers the variables in $M$ the state to be $\textbf{B}$ whenever 
+$P[y^L|M]\geq \frac{1}{2}$ and predicts $\textbf{R}$ otherwise._
+
+_Using the fact that for a random variable $z\sim N(\mu_z, \sigma_z)$ we have that $P[z\geq \mu_z]=\frac{1}{2}$ and 
+noticing that $M^c$ and $y^L|M$ are Normally distributed with means $\mathbb{E}[M^c]$ and $M+\mathbb{E}[M^c]$ 
+respectively, it is easy to see that  $P[y^L\geq 0 | M] \geq P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}$ whenever 
+$M+\mathbb{E}[M^c]\geq 0$. Similarly, whenever $M+\mathbb{E}[M^c]< 0$ we will have that 
+$P[y^L\geq 0 | M] < P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}$. Therefore, the optimal rule is to predict 
+$\mathbf{B}$ whenever $a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]$ and $\textbf{R}$ otherwise._
+
+With limited data, it is not guaranteed that the DM is able to perfectly learn the parameters required for the optimal rule. 
+However, the notion of optimal rules is useful for determining the scenarios in which we can expect polarization to arise. 
+The next section uses the notion of optimal rules to formalize the concept of polarization and to establish the 
+main prediction of the theory.
+
+## Polarization
+To study polarization, we take the definition of CITATION: two DMs are said to be 
+polarized if, when observing the same realization of observables, they predict the state to be different. That is, 
+they both observe the same draw of $x=(x_1, ..., x_N)$ but one DM predicts the state to be **B** and the other 
+predicts the state to be **R**.
+
+Within out framework---as in CITATION---polarization may persist even when DMs have 
+access to infinite data because they might be using different models of the world in order to make their predictions. 
+Nevertheless using different models does not necessarily imply that beliefs will always be polarized; it must also be 
+the case that the realization of observables falls in a *disagreement zone*. By disagreement zone, we simply 
+mean that if the two models that are being used for prediction are $M_1$ and $M_2$, the realization of the 
+observables $x=(x_1, ..., x_N)$ must be such that $M_1$ makes a different prediction from $M_2$. The following 
+2-dimensional example illustrates why the use of different models is necessary but not sufficient for polarization 
+to arise.
+
+#### Example
+_The optimal rule for the complex model coincides with the truth, it predicts **B** if $x_2+1\geq x_1$. 
+In addition, there are two simple models: one that considers only $x_1$---call it $M_1$---and one that considers 
+only $x_2$---call it $M_2$. The optimal rules for $M_1$ and $M_2$ are threshold rules that predict the state is **B** 
+if the value of $x_i$ is below (above) the threshold $t_i := \mu_{-i} + k$ and the predicted state is **R** otherwise.
+These prediction rules are illustrated in . As the figure makes clear, both rules agree on the predicted state for 
+some values of $(x_1, x_2)$ and disagree for other values. Theregions in which the rules make different predictions
+are the disagreement zones._
+
+For any pair of agents who observe the same realization of teh observables and use different models, it is possible
+to determine whether they will be polarized or not by looking at the predictions made by their models.
+
+# Experimental Design
 ```{r read-data-all}
@@
 
-## Including Plots
+Subjects for the experimet were recruited from the CESS lab subject pool at NYU. 
+We recruited `r data[, uniqueN(participant.code)]` undergraduate students who participated in this experiment in person at 
+the CESS lab. The experiment was coded using o-tree CITATION and it consisted of two parts: Part 1 was meant to expose the 
+subjects to the data generating process
+and allow them to learn how to predict; part 2 was meant to elicit the models of the world that subjects might have
+developed in part 1.
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Upon arrival subjects were randomly placed in pairs, however, they were not informed of this fact.
+Part 1 consited of 20 rounds in which subjects observed the realizations of all of the observables and were asked to predict
+the state. We had 5 observable variables called variable 1, 2, 3, 4 and 5. The variables were independent and Normally 
+distributed with means and variances as described in TABLE.
 
-\begin{equation}
-\lim_{x\to 0} \frac{\exp(x)-1}{x} = 1
-\end{equation}
+The state was determined by the following rule: $11x_1+6x_2+4.5x_3+2.4*x_4+k \geq 0$. The values of the coefficients were chosen
+so that all variables had a similar but unique level of informativeness when considered on their own. In this case, because
+all the variables are normally distributed, the information that they convey about the state is related to the
+variance of the modified normal $a_ix_i$. The constant $k$ was chosen so that the state was **B** in 50% of the cases. 
+The last column of TABLE shows the probability of the optimal prediction rule making a correct prediction when the model used
+considers only each variable by itself. This is the measure that we use to determine the informativeness of each variable and
+is reported in the last column of TABLE.
+
+| Variable | Mean | Variance | Modified Variable | Modified Variance | Informativeness |
+| -------- | ---- | -------- | ----------------- | ----------------- | --------------- |
+| $x_1$    | 0    |   1      | $11x_1$           |         11        |        .65      |
+| $x_2$    | -10  |   2      | $6x_2$            |         12        |        .67      |
+| $x_3$    | 5    |   3      | $4.5x_3$          |        13.5       |        .69      |
+| $x_4$    | -5   |   4      | $2.4*x_4$         |         12        |        .63      |
+| $x_5$    | 100  |   5      | $0x_5$            |         0         |         0       |
+
+In each round, subjects were asked to predict the state. They were given feedback on whether their prediction was correct or not.
+Both subjects within a pair were shown exactly the same realization of the variables and had to predict the same state. This allows
+us to determine whether the pair was polarized  or not for that particular observation.
+
+Part 2 started immediatly after part 1. In this part each subject was randomly allowed to use only a certain number of variables
+in order to make their prediction. Instead of showing them the realizations of all 5 variables, they were told that they can 
+disclose up to $K$ variables, where $K$ was drawn at random from the set $\{1, 2, ..., 5\}$ at every round and it was drawn 
+independently across subjects. We take the variables that they chose to reveal as the model that each subject using for predicting 
+the state and use that model to detrmine whether the theory would predict the pair to be polarized or not. In addition, we use 
+the random assignment of $K$ to investigate whether the subjects are using siumple or complex models. If allowing subjects to 
+disclose an additional variable improves their prediction accuracy, then it must be that they are in fact using the additional
+information for prediciton. If, on the other hand, they do not perform better when they have access to more information, it must be
+that although they are disclosing an additional variable, they do not correctly incorporate the infomation into their prediciton 
+rule. Thus, if we observe that the performance of our subjects does not improve when they are allowed to disclose $K+1$ variables, 
+relative to their performance when they were allowed to disclose only $K$ variables, their model of the world must be of size $K$
+or lower.
+
+# Data
+
+# Results
+
+# Conclusion
+
+# References
+
+# Appendix
 

### 2023-07-11 paper/paper.Rmd
writing predictions

@@
         citation_package: natbib
+bibliography: bibliography.bib
 fontsize: 12pt
@@
 A finite number of observable random variables, $X =(x_1, x_2, ..., x_N)$, determine a binary state, $y$. 
-The observable variables are independently distributed and each follows a Normal distribution with corresponding mean $\mu_i$ and standard deviation $\sigma_i$ with $i \in \{1, 2, ..., N\}$.
+The observable variables are independently distributed and each follows a Normal distribution with corresponding mean 
+$\mu_i$ and standard deviation $\sigma_i$ with $i \in \{1, 2, ..., N\}$.
 
@@
 A model of the world is said to be *simple* if it does not include all the relevant variables, 
-*i.e.* if $M\subset \mathcal{R}$. And it is complex if it includes all relevant variables, $M=\mathcal{R}$.
+*i.e.* if $M\subset \mathcal{R}$. And it is complex if it includes all relevant variables, $M=\mathcal{R}$. Thus, whenever 
+agents are using simple models, they are ignoring some of the relevant variables even if the information is available to them.
+In contrast, whenever agents are using complex models, they are using all the relevant information that is available to them, 
+so restricting the number of variables that is available to them will impair their ability to predict correctly. In the design
+of the experiment we will exploit these differences in order to determine whether subjects are using simple or complex models.
+
+**Prediction 1:** If subjects are using simple model with $K$ variables, then allowing them to disclose $K+1$ variables 
+will not increase the frequency with which they make correct predictions.
+
+**Prediction 1B:** If subjects are using complex models, then allowing them to disclose an additional variable will always 
+increase the frequency with which they make correct predictions.
 
@@
 ## Polarization
-To study polarization, we take the definition of CITATION: two DMs are said to be 
+To study polarization, we take the definition of @Haghtalab2021: two DMs are said to be 
 polarized if, when observing the same realization of observables, they predict the state to be different. That is, 
@@
 
-Within out framework---as in CITATION---polarization may persist even when DMs have 
+Within out framework---as in @Haghtalab2021---polarization may persist even when DMs have 
 access to infinite data because they might be using different models of the world in order to make their predictions. 
@@
 For any pair of agents who observe the same realization of teh observables and use different models, it is possible
-to determine whether they will be polarized or not by looking at the predictions made by their models.
+to determine whether they will be polarized or not by looking at the predictions made by their models. Notice that the 
+polarizing observations---those that fall into a polarization zone---are different for each pair of models. Also notice that 
+our definition of the polarization zone assumes that the DMs are using optimal rules. This is a very strong assumption 
+and will most likely not be the case that subjects in the experiment learn perfectly. However, the notion of optimal rules
+allows us to have a clean definition of polarization and to make predictions about when polarization will arise.
+
+**Prediction 2 :** Polarization will arise more often when the theory predicts so. That is, when two agents who have 
+different models of the world, are facing a realization of observables that falls into the polarization zone corresponding 
+to their models.
 
@@
 We recruited `r data[, uniqueN(participant.code)]` undergraduate students who participated in this experiment in person at 
-the CESS lab. The experiment was coded using o-tree CITATION and it consisted of two parts: Part 1 was meant to expose the 
+the CESS lab. The experiment was coded using o-tree @otree and it consisted of two parts: Part 1 was meant to expose the 
 subjects to the data generating process
@@
 relative to their performance when they were allowed to disclose only $K$ variables, their model of the world must be of size $K$
-or lower.
+or lower. This feature design allows us to test prediciton 1 and 1B and determine the number of variables that subjects are using.
 
@@
 
-# References
-
 # Appendix
-
-Inline code example: `r data[, uniqueN(participant.code)]`

### 2023-07-11 paper/paper.Rmd
writing learning results

@@
         citation_package: natbib
+        fig_width: 7
+        fig_height: 6
+        fig_caption: true
+        number_sections: false
+        keep_tex: true
+        extra_dependencies:
+            footmisc: ["bottom"] # footnote management
+            setspace: ["doublespacing"] # spacing of the paper
+            caption: ["normal"]
+            dsfont: null # indicator function 1
+            booktabs: null
+            makecell: null
+            hyperref: null
+
 bibliography: bibliography.bib
@@
 renv::snapshot()
-
+library("languageserver")
+library(reticulate)
 ```
@@
 
-## Optimal Rules
+### Optimal Rules
 The most accurate prediction rule for this setting will be one that predicts the state to be **B** whenever, 
@@
 
-## Polarization
+### Polarization
 To study polarization, we take the definition of @Haghtalab2021: two DMs are said to be 
@@
 
-#### Example
+##### Example
 _The optimal rule for the complex model coincides with the truth, it predicts **B** if $x_2+1\geq x_1$. 
@@
 data <- fread(here("data", "clean", "all.csv"))
+part1 <- fread(here("data", "clean", "part1.csv"))
+part2 <- fread(here("data", "clean", "part2.csv"))
+
 ```
@@
 
-Part 2 started immediatly after part 1. In this part each subject was randomly allowed to use only a certain number of variables
-in order to make their prediction. Instead of showing them the realizations of all 5 variables, they were told that they can 
-disclose up to $K$ variables, where $K$ was drawn at random from the set $\{1, 2, ..., 5\}$ at every round and it was drawn 
-independently across subjects. We take the variables that they chose to reveal as the model that each subject using for predicting 
-the state and use that model to detrmine whether the theory would predict the pair to be polarized or not. In addition, we use 
-the random assignment of $K$ to investigate whether the subjects are using siumple or complex models. If allowing subjects to 
-disclose an additional variable improves their prediction accuracy, then it must be that they are in fact using the additional
-information for prediciton. If, on the other hand, they do not perform better when they have access to more information, it must be
-that although they are disclosing an additional variable, they do not correctly incorporate the infomation into their prediciton 
-rule. Thus, if we observe that the performance of our subjects does not improve when they are allowed to disclose $K+1$ variables, 
+Part 2 started immediatly after part 1 and has 70 rounds. In this part each subject was randomly allowed to use only a certain 
+number of variables in order to make their prediction. Instead of showing them the realizations of all 5 variables, they were 
+told that they can disclose up to $K$ variables, where $K$ was drawn at random from the set $\{1, 2, ..., 5\}$. K was redrawn
+independently every round and across subjects. We use the random assignment of $K$ to investigate whether the subjects are using 
+siumple or complex models. If allowing subjects to disclose an additional variable improves their prediction accuracy, 
+then it must be that they are in fact using the additional information for prediciton. If, on the other hand, they do not perform 
+better when they have access to more information, it must be that although they are disclosing an additional variable, 
+they do not correctly incorporate the infomation into their prediciton rule. Thus, if we observe that the performance of our 
+subjects does not improve when they are allowed to disclose $K+1$ variables, 
 relative to their performance when they were allowed to disclose only $K$ variables, their model of the world must be of size $K$
@@
 
-# Data
+In order to explore the polarization prediciton, we use the fact that subjects are in fixed pairs throughout the experiment.
+Both subjects in a pair observe the same realization of all the variables in each round. Therefore, we can determine whether
+they are polarized or not by looking at whether they make the same prediction or not. We can then determine whether the theory 
+predicts that they should be polarized or not by looking at the polarization zone corresponding to the models that they are using.
+If the draw of the variables falls into the polarization zone, then the theory predicts that they should be polarized. The models
+that we use for this prediction are the ones that include the variables that they chose to reveal in that round. It could be that
+although they are revealing certain variables, they are not using them for prediction, however, we do not have a good way to 
+determine this. Instead, we use the variables that they reveal as a proxy for the variables that they are using for prediction and 
+ determine the polarization zones for those models.
 
 # Results
+
+In this section we present the results from the experiment. We start by exploring the general learning patterns of the subjects
+as well as the model choices. We then look at the effect of allowing subjects to disclose an additional variable and lastly we
+look at the polarization results.
+
+### Learning
+In the first part of the experiment, subjects were asked to predict the state given the realization of all the variables.
+Having access to all the information they could predict the state perfectly if they managed to learn the parameters of the
+hyperplane that determines the state. Figure \ref{learningAll} shows the share of correct predictions across rounds in part 1. 
+We see that subjects are able to learn how to predict the state at a rate that is higher than random. They initially predict
+the state corectly about 50% of the time, which is consistent with random guessing. The share of correct guesses increases to 
+`r 100*round(part1[subsession.round_number>15, mean(player.correct)], digits=2)`% by the end of part 1. And this last number is 
+significantly higher than 50% 
+(p-value = `r round(t.test(part1[subsession.round_number>10, player.correct], mu=0.5)$p.value, digits=4)`). 
+They continue to learn even after the first 20 rounds when we look at the rounds in which subjects had access to all the information.
+By the end of the experiment the rate of correct predictions is 
+`r 100*round(part2[subsession.round_number>30, mean(player.correct)], digits=2)`% which is still significantly higher than 50%
+(p-value = `r round(t.test(part2[subsession.round_number>30, player.correct], mu=0.5)$p.value, digits=4)`).
+
+```{r learningAll, out.width='50%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds'}
+knitr::include_graphics(here("computed_objects", "figures", "learning_all.png"))
+```
 

### 2023-07-11 paper/paper.tex
writing learning results

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2.5cm]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage[bottom]{footmisc}
\usepackage[doublespacing]{setspace}
\usepackage[normal]{caption}
\usepackage{dsfont}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{hyperref}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning from Data Through Models},
  pdfauthor={Alberto Bisin, Guillaume Frechette and Jimena Galindo},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Learning from Data Through Models}
\author{Alberto Bisin, Guillaume Frechette and Jimena Galindo}
\date{July 11, 2023}

\begin{document}
\maketitle
\begin{abstract}
TBW
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

TBW

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

TBW

\hypertarget{framework-and-predictions}{%
\section{Framework and Predictions}\label{framework-and-predictions}}

A finite number of observable random variables,
\(X =(x_1, x_2, ..., x_N)\), determine a binary state, \(y\). The
observable variables are independently distributed and each follows a
Normal distribution with corresponding mean \(\mu_i\) and standard
deviation \(\sigma_i\) with \(i \in \{1, 2, ..., N\}\).

The state is \emph{Red} (\textbf{R}) or \emph{Blue} (\textbf{B}) and is
determined in the following way: \begin{equation*}
y = \begin{cases}
\text{\textbf{B}} &\text{if $a_1x_1+...+a_Nx_N + K\geq0$}\\
\text{\textbf{R}} &\text{otherwise}
\end{cases}
\end{equation*}

An observable variable \(x_i\) is \emph{relevant} if the associated
coefficient, \(a_i\), is non-zero. Likewise, variable \(i\) is
\emph{irrelevant} if \(a_i = 0\). We call the set of relevant variables
\(\mathcal{R}\) and the set of irrelevant variables \(\mathcal{I}\).

In each period, the decision-maker (DM) observes the realization of all
observable variables and predicts the state. Their flow payoff in period
\(t\) is 1 if they predict the state correctly, and it is zero
otherwise. Denote the prediction by \(\hat{y}\).

In order to make a prediction, the DM uses a map
\(f: M_t\to \{\textbf{B}, \textbf{R}\}\)\\
where \(M_t \subseteq \{x_1, x_2, ..., x_N\}\) is the set of variables
that they choose to consider in period \(t\). Our main object of study
is the set of variables that the DM decides to consider and not the
properties of the mapping from the variables to the state. We refer to
the set of variables that are taken into consideration, \(M_t\), as the
DM's \emph{model of the world}, and we abstract away from how the
variables are mapped into a prediction of the state.

A model of the world is said to be \emph{simple} if it does not include
all the relevant variables, \emph{i.e.} if \(M\subset \mathcal{R}\). And
it is complex if it includes all relevant variables, \(M=\mathcal{R}\).
Thus, whenever agents are using simple models, they are ignoring some of
the relevant variables even if the information is available to them. In
contrast, whenever agents are using complex models, they are using all
the relevant information that is available to them, so restricting the
number of variables that is available to them will impair their ability
to predict correctly. In the design of the experiment we will exploit
these differences in order to determine whether subjects are using
simple or complex models.

\textbf{Prediction 1:} If subjects are using simple model with \(K\)
variables, then allowing them to disclose \(K+1\) variables will not
increase the frequency with which they make correct predictions.

\textbf{Prediction 1B:} If subjects are using complex models, then
allowing them to disclose an additional variable will always increase
the frequency with which they make correct predictions.

\hypertarget{optimal-rules}{%
\subsubsection{Optimal Rules}\label{optimal-rules}}

The most accurate prediction rule for this setting will be one that
predicts the state to be \textbf{B} whenever, given the observables,
\textbf{B} is more likely. And it predicts \textbf{R} whenever
\textbf{R} is more likely given the observables. Because we abstract
away from the procedure through which the rule is updated, we will make
predictions under the assumption that the DM has learned the parameters
perfectly. That is, they have seen enough information that, under an
updating procedure that converges to the truth---for example, correctly
specified regression model or Bayes rule---they would have converged to
the true parameters already.

We refer to a prediction rule as an \emph{Optimal Rule} if it is the
rule that a fully rational Decision maker would converge to with perfect
Bayesian learning. In this context, Bayesian learning corresponds to
using Bayes rule to learn the coefficients \(a = (a_1, a_2, ..., a_N)\)
as well as the constant \(K\). This means that in an \(N\) dimensional
space, the Decision Maker must learn \(N+1\) parameters which
corresponds to learning the halfspace in which the state is \textbf{B}.
The optimal rule for the complex model coincides with the true
data-generating process when learning is perfect.

Conditional on using a simple model, the optimal rule for that model is
the rule that, when restricted to the variables being considered, does
the best at predicting the state. In this case, the Decision Maker is
learning only the parameters that pertain the variables in the model.
Hence, for a model that considers \(K\) variables, the DM will have to
learn \(K+1\) parameters. The following proposition illustrates exactly
what the optimal prediction rules look like when the DM is able to
learn.

\hypertarget{proposition.}{%
\subparagraph{Proposition.}\label{proposition.}}

\emph{Fix any model that considers \(k\leq N\) variables and relabel the
variable so that \(x_1, ..., x_K\) are the variables that the model
considers and \(x_{k+1}, ..., x_N\) are the ones that the model does not
consider. Relabel the coefficients \(a = (a_1, ..., a_N)\) accordingly
as well.}

\hypertarget{proof.}{%
\subparagraph{proof.}\label{proof.}}

\emph{Let \(M:=a_1x_1+...+a_kx_k\) and
\(M^c = a_{k+1}x_{k+1}+...+a_Nx_N+k\) and define the latent variable
\(y^L := M + M^c\). The optimal prediction rule for a model that
considers the variables in \(M\) the state to be \(\textbf{B}\) whenever
\(P[y^L|M]\geq \frac{1}{2}\) and predicts \(\textbf{R}\) otherwise.}

\emph{Using the fact that for a random variable
\(z\sim N(\mu_z, \sigma_z)\) we have that \(P[z\geq \mu_z]=\frac{1}{2}\)
and noticing that \(M^c\) and \(y^L|M\) are Normally distributed with
means \(\mathbb{E}[M^c]\) and \(M+\mathbb{E}[M^c]\) respectively, it is
easy to see that
\(P[y^L\geq 0 | M] \geq P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}\)
whenever \(M+\mathbb{E}[M^c]\geq 0\). Similarly, whenever
\(M+\mathbb{E}[M^c]< 0\) we will have that
\(P[y^L\geq 0 | M] < P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}\).
Therefore, the optimal rule is to predict \(\mathbf{B}\) whenever
\(a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]\) and \(\textbf{R}\)
otherwise.}

With limited data, it is not guaranteed that the DM is able to perfectly
learn the parameters required for the optimal rule. However, the notion
of optimal rules is useful for determining the scenarios in which we can
expect polarization to arise. The next section uses the notion of
optimal rules to formalize the concept of polarization and to establish
the main prediction of the theory.

\hypertarget{polarization}{%
\subsubsection{Polarization}\label{polarization}}

To study polarization, we take the definition of \citet{Haghtalab2021}:
two DMs are said to be polarized if, when observing the same realization
of observables, they predict the state to be different. That is, they
both observe the same draw of \(x=(x_1, ..., x_N)\) but one DM predicts
the state to be \textbf{B} and the other predicts the state to be
\textbf{R}.

Within out framework---as in \citet{Haghtalab2021}---polarization may
persist even when DMs have access to infinite data because they might be
using different models of the world in order to make their predictions.
Nevertheless using different models does not necessarily imply that
beliefs will always be polarized; it must also be the case that the
realization of observables falls in a \emph{disagreement zone}. By
disagreement zone, we simply mean that if the two models that are being
used for prediction are \(M_1\) and \(M_2\), the realization of the
observables \(x=(x_1, ..., x_N)\) must be such that \(M_1\) makes a
different prediction from \(M_2\). The following 2-dimensional example
illustrates why the use of different models is necessary but not
sufficient for polarization to arise.

\hypertarget{example}{%
\subparagraph{Example}\label{example}}

\emph{The optimal rule for the complex model coincides with the truth,
it predicts \textbf{B} if \(x_2+1\geq x_1\). In addition, there are two
simple models: one that considers only \(x_1\)---call it \(M_1\)---and
one that considers only \(x_2\)---call it \(M_2\). The optimal rules for
\(M_1\) and \(M_2\) are threshold rules that predict the state is
\textbf{B} if the value of \(x_i\) is below (above) the threshold
\(t_i := \mu_{-i} + k\) and the predicted state is \textbf{R} otherwise.
These prediction rules are illustrated in . As the figure makes clear,
both rules agree on the predicted state for some values of
\((x_1, x_2)\) and disagree for other values. Theregions in which the
rules make different predictions are the disagreement zones.}

For any pair of agents who observe the same realization of teh
observables and use different models, it is possible to determine
whether they will be polarized or not by looking at the predictions made
by their models. Notice that the polarizing observations---those that
fall into a polarization zone---are different for each pair of models.
Also notice that our definition of the polarization zone assumes that
the DMs are using optimal rules. This is a very strong assumption and
will most likely not be the case that subjects in the experiment learn
perfectly. However, the notion of optimal rules allows us to have a
clean definition of polarization and to make predictions about when
polarization will arise.

\textbf{Prediction 2 :} Polarization will arise more often when the
theory predicts so. That is, when two agents who have different models
of the world, are facing a realization of observables that falls into
the polarization zone corresponding to their models.

\hypertarget{experimental-design}{%
\section{Experimental Design}\label{experimental-design}}

Subjects for the experimet were recruited from the CESS lab subject pool
at NYU. We recruited 30 undergraduate students who participated in this
experiment in person at the CESS lab. The experiment was coded using
o-tree \citet{otree} and it consisted of two parts: Part 1 was meant to
expose the subjects to the data generating process and allow them to
learn how to predict; part 2 was meant to elicit the models of the world
that subjects might have developed in part 1.

Upon arrival subjects were randomly placed in pairs, however, they were
not informed of this fact. Part 1 consited of 20 rounds in which
subjects observed the realizations of all of the observables and were
asked to predict the state. We had 5 observable variables called
variable 1, 2, 3, 4 and 5. The variables were independent and Normally
distributed with means and variances as described in TABLE.

The state was determined by the following rule:
\(11x_1+6x_2+4.5x_3+2.4*x_4+k \geq 0\). The values of the coefficients
were chosen so that all variables had a similar but unique level of
informativeness when considered on their own. In this case, because all
the variables are normally distributed, the information that they convey
about the state is related to the variance of the modified normal
\(a_ix_i\). The constant \(k\) was chosen so that the state was
\textbf{B} in 50\% of the cases. The last column of TABLE shows the
probability of the optimal prediction rule making a correct prediction
when the model used considers only each variable by itself. This is the
measure that we use to determine the informativeness of each variable
and is reported in the last column of TABLE.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1159}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0580}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1159}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2464}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2464}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2174}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Variance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modified Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modified Variance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Informativeness
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(x_1\) & 0 & 1 & \(11x_1\) & 11 & .65 \\
\(x_2\) & -10 & 2 & \(6x_2\) & 12 & .67 \\
\(x_3\) & 5 & 3 & \(4.5x_3\) & 13.5 & .69 \\
\(x_4\) & -5 & 4 & \(2.4*x_4\) & 12 & .63 \\
\(x_5\) & 100 & 5 & \(0x_5\) & 0 & 0 \\
\end{longtable}

In each round, subjects were asked to predict the state. They were given
feedback on whether their prediction was correct or not. Both subjects
within a pair were shown exactly the same realization of the variables
and had to predict the same state. This allows us to determine whether
the pair was polarized or not for that particular observation.

Part 2 started immediatly after part 1 and has 70 rounds. In this part
each subject was randomly allowed to use only a certain number of
variables in order to make their prediction. Instead of showing them the
realizations of all 5 variables, they were told that they can disclose
up to \(K\) variables, where \(K\) was drawn at random from the set
\(\{1, 2, ..., 5\}\). K was redrawn independently every round and across
subjects. We use the random assignment of \(K\) to investigate whether
the subjects are using siumple or complex models. If allowing subjects
to disclose an additional variable improves their prediction accuracy,
then it must be that they are in fact using the additional information
for prediciton. If, on the other hand, they do not perform better when
they have access to more information, it must be that although they are
disclosing an additional variable, they do not correctly incorporate the
infomation into their prediciton rule. Thus, if we observe that the
performance of our subjects does not improve when they are allowed to
disclose \(K+1\) variables, relative to their performance when they were
allowed to disclose only \(K\) variables, their model of the world must
be of size \(K\) or lower. This feature design allows us to test
prediciton 1 and 1B and determine the number of variables that subjects
are using.

In order to explore the polarization prediciton, we use the fact that
subjects are in fixed pairs throughout the experiment. Both subjects in
a pair observe the same realization of all the variables in each round.
Therefore, we can determine whether they are polarized or not by looking
at whether they make the same prediction or not. We can then determine
whether the theory predicts that they should be polarized or not by
looking at the polarization zone corresponding to the models that they
are using. If the draw of the variables falls into the polarization
zone, then the theory predicts that they should be polarized. The models
that we use for this prediction are the ones that include the variables
that they chose to reveal in that round. It could be that although they
are revealing certain variables, they are not using them for prediction,
however, we do not have a good way to determine this. Instead, we use
the variables that they reveal as a proxy for the variables that they
are using for prediction and determine the polarization zones for those
models.

\hypertarget{results}{%
\section{Results}\label{results}}

In this section we present the results from the experiment. We start by
exploring the general learning patterns of the subjects as well as the
model choices. We then look at the effect of allowing subjects to
disclose an additional variable and lastly we look at the polarization
results.

\hypertarget{learning}{%
\subsubsection{Learning}\label{learning}}

In the first part of the experiment, subjects were asked to predict the
state given the realization of all the variables. Having access to all
the information they could predict the state perfectly if they managed
to learn the parameters of the hyperplane that determines the state.
Figure \ref{learningAll} shows the share of correct predictions across
rounds in part 1. We see that subjects are able to learn how to predict
the state at a rate that is higher than random. They initially predict
the state corectly about 50\% of the time, which is consistent with
random guessing. The share of correct guesses increases to 59\% by the
end of part 1. And this last number is significantly higher than 50\%
(p-value = \ensuremath{3\times 10^{-4}}). They continue to learn even
after the first 20 rounds when we look at the rounds in which subjects
had access to all the information. By the end of the experiment the rate
of correct predictions is 59\% which is still significantly higher than
50\% (p-value = 0)

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/learning_all}

}

\caption{\label{learningAll} Share of correct predictions across rounds}\label{fig:learningAll}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\renewcommand\refname{Appendix}
  \bibliography{bibliography.bib}

\end{document}

### 2023-07-11 analysis/learning_plots.py

@@ -23,7 +23,7 @@ sns.lmplot(data=all_rounds[all_rounds['round_number_modif']<21], x="subsession.r
             x_bins=5, aspect=1.5, height=5)
 # add a line at .5 that indicates the random guess level 
 plt.axhline(.5, 0, 1, color = 'grey')
-
+plt.tight_layout()
 plt.xlabel('round number')
 plt.ylabel('share correct guesses')
 
@@ -35,7 +35,7 @@ plt.savefig('computed_objects/figures/learning_part1.png')
 # truncate the data at round 60 since there are only few observations after that
 sns.lmplot(data=all_rounds[(all_rounds['player.number_variables']>3) & (all_rounds['round_number_modif']<61)], x="round_number_modif", y="player.correct", 
             x_bins=10, aspect=1.5, height=5)
-
+plt.tight_layout()
 plt.xlabel('round number')
 plt.ylabel('share correct')
 plt.axhline(.5, 0, 1, color = 'grey')

### 2023-07-11 paper/paper.Rmd
edit framework writing

@@
 The observable variables are independently distributed and each follows a Normal distribution with corresponding mean 
-$\mu_i$ and standard deviation $\sigma_i$ with $i \in \{1, 2, ..., N\}$.
+$\mu_i$ and standard deviation $\sigma_i$. The state is \emph{Red} (**R**) or \emph{Blue} (**B**) and is determined in 
+the following way:
 
-The state is \emph{Red} (**R**) or \emph{Blue} (**B**) and is determined in the following way:
 \begin{equation*}
@@
 An observable variable $x_i$ is *relevant* if the associated coefficient, $a_i$, is non-zero. 
-Likewise, variable $i$ is *irrelevant* if $a_i = 0$. 
-We call the set of relevant variables $\mathcal{R}$ and the set of irrelevant variables $\mathcal{I}$.
+Likewise, variable $i$ is *irrelevant* if $a_i = 0$. We call the set of relevant variables $\mathcal{R}$ and the set of 
+irrelevant variables $\mathcal{I}$.
 
-In each period, the decision-maker (DM) observes the realization of all observable variables and predicts the state. 
-Their flow payoff in period $t$ is 1 if they predict the state correctly, and it is zero otherwise. 
-Denote the prediction by $\hat{y}$.
+In each period, an agent observes the realization of all observable variables and predicts the state. Their flow payoff in 
+period $t$ is 1 if they predict the state correctly, and it is zero otherwise. Denote the prediction by $\hat{y}$.
 
-In order to make a prediction, the DM uses a map $f: M_t\to \{\textbf{B}, \textbf{R}\}$  
-where $M_t \subseteq \{x_1, x_2, ..., x_N\}$ is the set of variables that they choose to consider in period $t$. 
-Our main object of study is the set of variables that the DM decides to consider and not the properties of the 
-mapping from the variables to the state. We refer to the set of variables that are taken into consideration, $M_t$, 
-as the DM's *model of the world*, and we abstract away from how the variables are mapped into a prediction of 
-the state.
+In order to make a prediction, the agent uses a map $M_t\to \{\textbf{B}, \textbf{R}\}$  where 
+$M_t \subseteq \{x_1, x_2, ..., x_N\}$ is the set of variables that they choose to consider in period $t$. 
+Our main object of study is the set of variables that the agent decides to consider in a given period, and the procedure by 
+which the maping or the set are updated. We refer to the set of variables that are taken into consideration, $M_t$, as the agent's 
+*mental model*, and assume that the mapping is linear.
 
-A model of the world is said to be *simple* if it does not include all the relevant variables, 
-*i.e.* if $M\subset \mathcal{R}$. And it is complex if it includes all relevant variables, $M=\mathcal{R}$. Thus, whenever 
-agents are using simple models, they are ignoring some of the relevant variables even if the information is available to them.
-In contrast, whenever agents are using complex models, they are using all the relevant information that is available to them, 
-so restricting the number of variables that is available to them will impair their ability to predict correctly. In the design
-of the experiment we will exploit these differences in order to determine whether subjects are using simple or complex models.
+A mental model is *simple* if it does not include all the relevant variables, 
+*i.e.* if $M\subset \mathcal{R}$. And it is *complex* if it includes all relevant variables, $M=\mathcal{R}$. In 
+addition, mental models can be ranked in terms of their simplicity by the number of relevant variables that they consider.
 
-**Prediction 1:** If subjects are using simple model with $K$ variables, then allowing them to disclose $K+1$ variables 
+Whenever agents use simple models, they ignore some of the relevant information that is available to them.
+In contrast, whenever agents use complex models, they use all the relevant information available. This means that when agents 
+with complex mental models are restricted in terms of how many variables they can consider, their ability
+to predict the state will be impaired. On the contrary, restricting the number of variables that an agent with a simple model can 
+use, does not necessarily affect their ability to predict the state. In particular, if they are considering fewer variables than 
+what the restriciton allows, their predictions should be as accurate as without the restriciton. These two effects are captured in 
+predictions 1 and 1B respectively.
+
+**Prediction 1:** If subjects use complex models, then allowing them to use an additional variable will always 
+increase the frequency with which they make correct predictions.
+
+**Prediction 1B:** If subjects use a simple mental model that considers $K$ variables, allowing them to use $K+1$ variables 
 will not increase the frequency with which they make correct predictions.
 
-**Prediction 1B:** If subjects are using complex models, then allowing them to disclose an additional variable will always 
-increase the frequency with which they make correct predictions.
+### Optimal Rules
+Conditional of using a particular mental model, the most accurate prediction rule that the agent can use is one that 
+predicts the state to be **B** whenever, given the observables under consideration, **B** is more likely. 
+And it predicts **R** whenever **R** is more likely. Because we abstract away from the procedure through which the rule is updated, 
+we will make predictions under the assumption that the agent has learned the parameters perfectly. That is, 
+they have seen enough information that, under an updating procedure that converges to the truth---for example, 
+correctly specified regression model or Bayes rule---they would have converged to the true parameters already. This provides an 
+upper bound for how well any model can perform.
 
-### Optimal Rules
-The most accurate prediction rule for this setting will be one that predicts the state to be **B** whenever, 
-given the observables, **B** is more likely. And it predicts **R** whenever **R** is more likely 
-given the observables. Because we abstract away from the procedure through which the rule is updated, 
-we will make predictions under the assumption that the DM has learned the parameters perfectly. That is, 
-they have seen enough information that, under an updating procedure that converges to the truth---for example, 
-correctly specified regression model or Bayes rule---they would have converged to the true parameters already.
-
-We refer to a prediction rule as an *Optimal Rule* if it is the rule that a fully rational Decision maker would 
+We refer to a prediction rule as an *Optimal Rule* if it is the rule that a fully rational agent would 
 converge to with perfect Bayesian learning. In this context, Bayesian learning corresponds to using Bayes rule to 
 learn the coefficients $a = (a_1, a_2, ..., a_N)$ as well as the constant $K$. This means that in an $N$ dimensional 
-space, the Decision Maker must learn $N+1$ parameters which corresponds to learning the halfspace in which the state 
+space, the agent must learn $N+1$ parameters which corresponds to learning the halfspace in which the state 
 is **B**. The optimal rule for the complex model coincides with the true data-generating process when learning 
-is perfect.
+is perfect and the parameters are identified.
 
-Conditional on using a simple model, the optimal rule for that model is the rule that, when restricted to the 
-variables being considered, does the best at predicting the state. In this case, the Decision Maker is learning only 
-the parameters that pertain the variables in the model. Hence, for a model that considers $K$ variables, the DM will 
-have to learn $K+1$ parameters. The following proposition illustrates exactly what the optimal prediction rules look 
-like when the DM is able to learn.
+An optimal rule conditional on model $M$ is the rule that, when restricted to the 
+variables being considered, does the best at predicting the state. In this case, the agent is learning only 
+the parameters that pertain the variables in $M$ plus a constant term. The following proposition illustrates exactly 
+what the optimal prediction rules look like when the agent is able to learn perfectly.
 
 ##### Proposition.
-_Fix any model that considers $k\leq N$ variables and relabel the variable so that $x_1, ..., x_K$ are the variables 
-that the model considers and $x_{k+1}, ..., x_N$ are the ones that the model does not consider. Relabel the 
+_Fix any model that considers $m\leq N$ variables and relabel the variable so that $x_1, ..., x_m$ are the variables 
+that the model considers and $x_{m+1}, ..., x_N$ are the ones that the model does not consider. Relabel the 
 coefficients $a = (a_1, ..., a_N)$ accordingly as well._
@@
 ##### proof.
-_Let $M:=a_1x_1+...+a_kx_k$ and $M^c = a_{k+1}x_{k+1}+...+a_Nx_N+k$ and define the latent variable $y^L := M +  M^c$. 
-The optimal prediction rule for a model that considers the variables in $M$ the state to be $\textbf{B}$ whenever 
+_Let $M:=a_1x_1+...+a_mx_m$ and $M^c = a_{m+1}x_{m+1}+...+a_Nx_N+k$ and define the latent variable $y^L := M +  M^c$. 
+The optimal prediction rule for model $M$ predicts the state to be $\textbf{B}$ whenever 
 $P[y^L|M]\geq \frac{1}{2}$ and predicts $\textbf{R}$ otherwise._
@@
 
-With limited data, it is not guaranteed that the DM is able to perfectly learn the parameters required for the optimal rule. 
-However, the notion of optimal rules is useful for determining the scenarios in which we can expect polarization to arise. 
+With limited data, it is not guaranteed that the agent is able to perfectly learn the parameters required for the optimal rule. 
+However, the notion of optimal rules is useful for determining a benchmark for how well each model can perform, as well as 
+the scenarios in which we can expect polarization to arise. 
 The next section uses the notion of optimal rules to formalize the concept of polarization and to establish the 
@@
 ### Polarization
-To study polarization, we take the definition of @Haghtalab2021: two DMs are said to be 
+We take the definition of polarization from @Haghtalab2021: two agents are said to be 
 polarized if, when observing the same realization of observables, they predict the state to be different. That is, 
-they both observe the same draw of $x=(x_1, ..., x_N)$ but one DM predicts the state to be **B** and the other 
+they both observe the same draw of $x=(x_1, ..., x_N)$ but one agent predicts the state to be **B** and the other 
 predicts the state to be **R**.
 
-Within out framework---as in @Haghtalab2021---polarization may persist even when DMs have 
-access to infinite data because they might be using different models of the world in order to make their predictions. 
-Nevertheless using different models does not necessarily imply that beliefs will always be polarized; it must also be 
-the case that the realization of observables falls in a *disagreement zone*. By disagreement zone, we simply 
-mean that if the two models that are being used for prediction are $M_1$ and $M_2$, the realization of the 
-observables $x=(x_1, ..., x_N)$ must be such that $M_1$ makes a different prediction from $M_2$. The following 
+Within out framework---as in @Haghtalab2021---polarization may persist even when agents have 
+access to infinite data because they might be using different models of the world. 
+Nevertheless using different models does not necessarily imply that beliefs will always be polarized, it must also be 
+the case that the realization of observables falls in a polarization zone. By *polarization zone*, we  
+mean that if the two models being used for prediction are $M_1$ and $M_2$, the realization of the 
+observables $x=(x_1, ..., x_N)$ must be such that the optimal rule for $M_1$ makes a different prediction from 
+that of $M_2$. The following 
 2-dimensional example illustrates why the use of different models is necessary but not sufficient for polarization 
@@
 ##### Example
-_The optimal rule for the complex model coincides with the truth, it predicts **B** if $x_2+1\geq x_1$. 
+_The optimal rule for the complex model coincides with the truth, it predicts **B** if $x_2-1\geq x_1$. 
 In addition, there are two simple models: one that considers only $x_1$---call it $M_1$---and one that considers 
@@
 if the value of $x_i$ is below (above) the threshold $t_i := \mu_{-i} + k$ and the predicted state is **R** otherwise.
-These prediction rules are illustrated in . As the figure makes clear, both rules agree on the predicted state for 
-some values of $(x_1, x_2)$ and disagree for other values. Theregions in which the rules make different predictions
+These prediction rules are illustrated in . As Figure \ref{example} makes clear, both rules agree on the predicted state for 
+some values of $(x_1, x_2)$ and disagree for other values. The regions in which the rules make different predictions
 are the disagreement zones._
 
-For any pair of agents who observe the same realization of teh observables and use different models, it is possible
-to determine whether they will be polarized or not by looking at the predictions made by their models. Notice that the 
+```{r example, out.width='50%', fig.align='center', fig.cap='\\label{example} Polarization regions for two single variable models'}
+knitr::include_graphics(here("diagram_polariazation_2variables.png"))
+```
+
+For any pair of agents who observe the same realization of the observables and use different models, it is possible
+to determine whether they will be polarized or not by looking at the predictions made by the optimal rule corresponding to their 
+models. Notice that the 
 polarizing observations---those that fall into a polarization zone---are different for each pair of models. Also notice that 
-our definition of the polarization zone assumes that the DMs are using optimal rules. This is a very strong assumption 
-and will most likely not be the case that subjects in the experiment learn perfectly. However, the notion of optimal rules
-allows us to have a clean definition of polarization and to make predictions about when polarization will arise.
+our definition of the polarization zone assumes that the agents are using optimal rules. This is a very strong assumption 
+and will most likely not be the case that subjects in the experiment learn perfectly. This is a theoretical construct that allows
+us to formalize Prediction 2.
 
@@
 Subjects for the experimet were recruited from the CESS lab subject pool at NYU. 
-We recruited `r data[, uniqueN(participant.code)]` undergraduate students who participated in this experiment in person at 
-the CESS lab. The experiment was coded using o-tree @otree and it consisted of two parts: Part 1 was meant to expose the 
-subjects to the data generating process
+We recruited `r data[, uniqueN(participant.code)]` undergraduate students who participated in this experiment in person 
+durinf the summer and fall of 2023. The experiment was coded using o-tree @otree and it consisted of two parts:
+Part 1 was meant to expose the subjects to the data generating process
 and allow them to learn how to predict; part 2 was meant to elicit the models of the world that subjects might have
@@
 
-Upon arrival subjects were randomly placed in pairs, however, they were not informed of this fact.
 Part 1 consited of 20 rounds in which subjects observed the realizations of all of the observables and were asked to predict

### 2023-07-11 paper/paper.tex
edit framework writing

@@
 Normal distribution with corresponding mean \(\mu_i\) and standard
-deviation \(\sigma_i\) with \(i \in \{1, 2, ..., N\}\).
+deviation \(\sigma_i\). The state is \emph{Red} (\textbf{R}) or
+\emph{Blue} (\textbf{B}) and is determined in the following way:
 
-The state is \emph{Red} (\textbf{R}) or \emph{Blue} (\textbf{B}) and is
-determined in the following way: \begin{equation*}
+\begin{equation*}
 y = \begin{cases}
@@
 
-In each period, the decision-maker (DM) observes the realization of all
-observable variables and predicts the state. Their flow payoff in period
-\(t\) is 1 if they predict the state correctly, and it is zero
-otherwise. Denote the prediction by \(\hat{y}\).
+In each period, an agent observes the realization of all observable
+variables and predicts the state. Their flow payoff in period \(t\) is 1
+if they predict the state correctly, and it is zero otherwise. Denote
+the prediction by \(\hat{y}\).
 
-In order to make a prediction, the DM uses a map
-\(f: M_t\to \{\textbf{B}, \textbf{R}\}\)\\
-where \(M_t \subseteq \{x_1, x_2, ..., x_N\}\) is the set of variables
-that they choose to consider in period \(t\). Our main object of study
-is the set of variables that the DM decides to consider and not the
-properties of the mapping from the variables to the state. We refer to
-the set of variables that are taken into consideration, \(M_t\), as the
-DM's \emph{model of the world}, and we abstract away from how the
-variables are mapped into a prediction of the state.
+In order to make a prediction, the agent uses a map
+\(M_t\to \{\textbf{B}, \textbf{R}\}\) where
+\(M_t \subseteq \{x_1, x_2, ..., x_N\}\) is the set of variables that
+they choose to consider in period \(t\). Our main object of study is the
+set of variables that the agent decides to consider in a given period,
+and the procedure by which the maping or the set are updated. We refer
+to the set of variables that are taken into consideration, \(M_t\), as
+the agent's \emph{mental model}, and assume that the mapping is linear.
 
-A model of the world is said to be \emph{simple} if it does not include
-all the relevant variables, \emph{i.e.} if \(M\subset \mathcal{R}\). And
-it is complex if it includes all relevant variables, \(M=\mathcal{R}\).
-Thus, whenever agents are using simple models, they are ignoring some of
-the relevant variables even if the information is available to them. In
-contrast, whenever agents are using complex models, they are using all
-the relevant information that is available to them, so restricting the
-number of variables that is available to them will impair their ability
-to predict correctly. In the design of the experiment we will exploit
-these differences in order to determine whether subjects are using
-simple or complex models.
+A mental model is \emph{simple} if it does not include all the relevant
+variables, \emph{i.e.} if \(M\subset \mathcal{R}\). And it is
+\emph{complex} if it includes all relevant variables, \(M=\mathcal{R}\).
+In addition, mental models can be ranked in terms of their simplicity by
+the number of relevant variables that they consider.
 
-\textbf{Prediction 1:} If subjects are using simple model with \(K\)
-variables, then allowing them to disclose \(K+1\) variables will not
-increase the frequency with which they make correct predictions.
+Whenever agents use simple models, they ignore some of the relevant
+information that is available to them. In contrast, whenever agents use
+complex models, they use all the relevant information available. This
+means that when agents with complex mental models are restricted in
+terms of how many variables they can consider, their ability to predict
+the state will be impaired. On the contrary, restricting the number of
+variables that an agent with a simple model can use, does not
+necessarily affect their ability to predict the state. In particular, if
+they are considering fewer variables than what the restriciton allows,
+their predictions should be as accurate as without the restriciton.
+These two effects are captured in predictions 1 and 1B respectively.
 
-\textbf{Prediction 1B:} If subjects are using complex models, then
-allowing them to disclose an additional variable will always increase
-the frequency with which they make correct predictions.
+\textbf{Prediction 1:} If subjects use complex models, then allowing
+them to use an additional variable will always increase the frequency
+with which they make correct predictions.
+
+\textbf{Prediction 1B:} If subjects use a simple mental model that
+considers \(K\) variables, allowing them to use \(K+1\) variables will
+not increase the frequency with which they make correct predictions.
 
@@
 
-The most accurate prediction rule for this setting will be one that
-predicts the state to be \textbf{B} whenever, given the observables,
+Conditional of using a particular mental model, the most accurate
+prediction rule that the agent can use is one that predicts the state to
+be \textbf{B} whenever, given the observables under consideration,
 \textbf{B} is more likely. And it predicts \textbf{R} whenever
-\textbf{R} is more likely given the observables. Because we abstract
-away from the procedure through which the rule is updated, we will make
-predictions under the assumption that the DM has learned the parameters
-perfectly. That is, they have seen enough information that, under an
-updating procedure that converges to the truth---for example, correctly
-specified regression model or Bayes rule---they would have converged to
-the true parameters already.
+\textbf{R} is more likely. Because we abstract away from the procedure
+through which the rule is updated, we will make predictions under the
+assumption that the agent has learned the parameters perfectly. That is,
+they have seen enough information that, under an updating procedure that
+converges to the truth---for example, correctly specified regression
+model or Bayes rule---they would have converged to the true parameters
+already. This provides an upper bound for how well any model can
+perform.
 
 We refer to a prediction rule as an \emph{Optimal Rule} if it is the
-rule that a fully rational Decision maker would converge to with perfect
-Bayesian learning. In this context, Bayesian learning corresponds to
-using Bayes rule to learn the coefficients \(a = (a_1, a_2, ..., a_N)\)
-as well as the constant \(K\). This means that in an \(N\) dimensional
-space, the Decision Maker must learn \(N+1\) parameters which
-corresponds to learning the halfspace in which the state is \textbf{B}.
-The optimal rule for the complex model coincides with the true
-data-generating process when learning is perfect.
+rule that a fully rational agent would converge to with perfect Bayesian
+learning. In this context, Bayesian learning corresponds to using Bayes
+rule to learn the coefficients \(a = (a_1, a_2, ..., a_N)\) as well as
+the constant \(K\). This means that in an \(N\) dimensional space, the
+agent must learn \(N+1\) parameters which corresponds to learning the
+halfspace in which the state is \textbf{B}. The optimal rule for the
+complex model coincides with the true data-generating process when
+learning is perfect and the parameters are identified.
 
-Conditional on using a simple model, the optimal rule for that model is
-the rule that, when restricted to the variables being considered, does
-the best at predicting the state. In this case, the Decision Maker is
-learning only the parameters that pertain the variables in the model.
-Hence, for a model that considers \(K\) variables, the DM will have to
-learn \(K+1\) parameters. The following proposition illustrates exactly
-what the optimal prediction rules look like when the DM is able to
-learn.
+An optimal rule conditional on model \(M\) is the rule that, when
+restricted to the variables being considered, does the best at
+predicting the state. In this case, the agent is learning only the
+parameters that pertain the variables in \(M\) plus a constant term. The
+following proposition illustrates exactly what the optimal prediction
+rules look like when the agent is able to learn perfectly.
 
@@
 
-\emph{Fix any model that considers \(k\leq N\) variables and relabel the
-variable so that \(x_1, ..., x_K\) are the variables that the model
-considers and \(x_{k+1}, ..., x_N\) are the ones that the model does not
+\emph{Fix any model that considers \(m\leq N\) variables and relabel the
+variable so that \(x_1, ..., x_m\) are the variables that the model
+considers and \(x_{m+1}, ..., x_N\) are the ones that the model does not
 consider. Relabel the coefficients \(a = (a_1, ..., a_N)\) accordingly
@@
 
-\emph{Let \(M:=a_1x_1+...+a_kx_k\) and
-\(M^c = a_{k+1}x_{k+1}+...+a_Nx_N+k\) and define the latent variable
-\(y^L := M + M^c\). The optimal prediction rule for a model that
-considers the variables in \(M\) the state to be \(\textbf{B}\) whenever
-\(P[y^L|M]\geq \frac{1}{2}\) and predicts \(\textbf{R}\) otherwise.}
+\emph{Let \(M:=a_1x_1+...+a_mx_m\) and
+\(M^c = a_{m+1}x_{m+1}+...+a_Nx_N+k\) and define the latent variable
+\(y^L := M + M^c\). The optimal prediction rule for model \(M\) predicts
+the state to be \(\textbf{B}\) whenever \(P[y^L|M]\geq \frac{1}{2}\) and
+predicts \(\textbf{R}\) otherwise.}
 
@@
 
-With limited data, it is not guaranteed that the DM is able to perfectly
-learn the parameters required for the optimal rule. However, the notion
-of optimal rules is useful for determining the scenarios in which we can
-expect polarization to arise. The next section uses the notion of
+With limited data, it is not guaranteed that the agent is able to
+perfectly learn the parameters required for the optimal rule. However,
+the notion of optimal rules is useful for determining a benchmark for
+how well each model can perform, as well as the scenarios in which we
+can expect polarization to arise. The next section uses the notion of
 optimal rules to formalize the concept of polarization and to establish
@@
 
-To study polarization, we take the definition of \citet{Haghtalab2021}:
-two DMs are said to be polarized if, when observing the same realization
+We take the definition of polarization from \citet{Haghtalab2021}: two
+agents are said to be polarized if, when observing the same realization
 of observables, they predict the state to be different. That is, they
-both observe the same draw of \(x=(x_1, ..., x_N)\) but one DM predicts
-the state to be \textbf{B} and the other predicts the state to be
-\textbf{R}.
+both observe the same draw of \(x=(x_1, ..., x_N)\) but one agent
+predicts the state to be \textbf{B} and the other predicts the state to
+be \textbf{R}.
 
 Within out framework---as in \citet{Haghtalab2021}---polarization may
-persist even when DMs have access to infinite data because they might be
-using different models of the world in order to make their predictions.
-Nevertheless using different models does not necessarily imply that
-beliefs will always be polarized; it must also be the case that the
-realization of observables falls in a \emph{disagreement zone}. By
-disagreement zone, we simply mean that if the two models that are being
-used for prediction are \(M_1\) and \(M_2\), the realization of the
-observables \(x=(x_1, ..., x_N)\) must be such that \(M_1\) makes a
-different prediction from \(M_2\). The following 2-dimensional example
-illustrates why the use of different models is necessary but not
-sufficient for polarization to arise.
+persist even when agents have access to infinite data because they might
+be using different models of the world. Nevertheless using different
+models does not necessarily imply that beliefs will always be polarized,
+it must also be the case that the realization of observables falls in a
+\emph{disagreement zone}. By disagreement zone, we simply mean that if
+the two models that are being used for prediction are \(M_1\) and
+\(M_2\), the realization of the observables \(x=(x_1, ..., x_N)\) must
+be such that \(M_1\) makes a different prediction from \(M_2\). The
+following 2-dimensional example illustrates why the use of different
+models is necessary but not sufficient for polarization to arise.
 
@@
 \(t_i := \mu_{-i} + k\) and the predicted state is \textbf{R} otherwise.
-These prediction rules are illustrated in . As the figure makes clear,
-both rules agree on the predicted state for some values of
-\((x_1, x_2)\) and disagree for other values. Theregions in which the
+These prediction rules are illustrated in . As Figure \ref{example}
+makes clear, both rules agree on the predicted state for some values of
+\((x_1, x_2)\) and disagree for other values. The regions in which the
 rules make different predictions are the disagreement zones.}
 
-For any pair of agents who observe the same realization of teh
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{../diagram_polariazation_2variables}
+
+}
+
+\caption{\label{example} Polarization regions for two single variable models}\label{fig:example}
+\end{figure}
+
+For any pair of agents who observe the same realization of the
 observables and use different models, it is possible to determine
 whether they will be polarized or not by looking at the predictions made
-by their models. Notice that the polarizing observations---those that
-fall into a polarization zone---are different for each pair of models.
-Also notice that our definition of the polarization zone assumes that
-the DMs are using optimal rules. This is a very strong assumption and
-will most likely not be the case that subjects in the experiment learn
-perfectly. However, the notion of optimal rules allows us to have a
-clean definition of polarization and to make predictions about when
-polarization will arise.
+by the optimal rule corresponding to their models. Notice that the
+polarizing observations---those that fall into a polarization zone---are
+different for each pair of models. Also notice that our definition of
+the polarization zone assumes that the agents are using optimal rules.
+This is a very strong assumption and will most likely not be the case
+that subjects in the experiment learn perfectly. This is a theoretical
+construct that allows us to formalize Prediction 2.
 
@@
 at NYU. We recruited 30 undergraduate students who participated in this
-experiment in person at the CESS lab. The experiment was coded using
-o-tree \citet{otree} and it consisted of two parts: Part 1 was meant to
-expose the subjects to the data generating process and allow them to
-learn how to predict; part 2 was meant to elicit the models of the world
-that subjects might have developed in part 1.
+experiment in person durinf the summer and fall of 2023. The experiment
+was coded using o-tree \citet{otree} and it consisted of two parts: Part
+1 was meant to expose the subjects to the data generating process and
+allow them to learn how to predict; part 2 was meant to elicit the
+models of the world that subjects might have developed in part 1.
 
-Upon arrival subjects were randomly placed in pairs, however, they were
-not informed of this fact. Part 1 consited of 20 rounds in which
-subjects observed the realizations of all of the observables and were
-asked to predict the state. We had 5 observable variables called
-variable 1, 2, 3, 4 and 5. The variables were independent and Normally
-distributed with means and variances as described in TABLE.
+Part 1 consited of 20 rounds in which subjects observed the realizations
+of all of the observables and were asked to predict the state. We had 5
+observable variables called variable 1, 2, 3, 4 and 5. The variables
+were independent and Normally distributed with means and variances as
+described in TABLE.
 
@@
 of correct predictions is 59\% which is still significantly higher than
-50\% (p-value = 0)
+50\% (p-value = 0).
 

### 2023-07-13 paper/paper.Rmd
regressions in r

@@
 abstract: "TBW"
+#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
 output: 
@@
         fig_caption: true
-        number_sections: false
+        number_sections: true
         keep_tex: true
@@
             hyperref: null
+        includes:
+          in_header: extra_header.tex
 
@@
 fontsize: 12pt
-geometry:
-- margin=2.5cm
-hyperref:
-- linestretch=2.0
-documentclass: article
+geometry: margin=2.5cm
 ---
@@
 library(here)
+library(huxtable)
+library(broom)
 renv::snapshot()
@@
 library(reticulate)
+
+# Format the numbers in the inline code output
+inline_hook <- function(x){
+  if(is.numeric(x)) {
+
+    # integers get , to separate thousands
+    if(abs(x - round(x)) < .Machine$double.eps){
+      formatted <- format(x, big.mark = ",", scientific = FALSE)
+    } else {
+    # consistent rounding of non-integers
+      formatted <- format(x, digits = 2, nsmall = 0, scientific = FALSE)
+    }
+
+  } else{
+
+    # render text in bold (commented out)
+    if(is.character(x)){
+      # formatted <- paste0("**", x, "**")
+      formatted <- x
+    } else{
+      formatted <- x
+    }
+
+  }
+
+  formatted
+}
+
+knitr::knit_hooks$set(inline = inline_hook)
 ```
@@
 # Introduction
-TBW
 
 # Literature Review
-TBW
 
@@
 
-### Optimal Rules
+## Optimal Rules
 Conditional of using a particular mental model, the most accurate prediction rule that the agent can use is one that 
@@
 
-##### Proposition.
-_Fix any model that considers $m\leq N$ variables and relabel the variable so that $x_1, ..., x_m$ are the variables 
+\begin{proposition}
+Fix any model that considers $m\leq N$ variables and relabel the variable so that $x_1, ..., x_m$ are the variables 
 that the model considers and $x_{m+1}, ..., x_N$ are the ones that the model does not consider. Relabel the 
-coefficients $a = (a_1, ..., a_N)$ accordingly as well._
+coefficients $a = (a_1, ..., a_N)$ accordingly as well. 
+\end{proposition}
+*Proof*.
+Let $M:=a_1x_1+...+a_mx_m$ and $M^c = a_{m+1}x_{m+1}+...+a_Nx_N+k$ and define the latent variable $y^L := M +  M^c$. 
+The optimal prediction rule for model $M$ predicts the state to be $\textbf{B}$ whenever 
+$P[y^L|M]\geq \frac{1}{2}$ and predicts $\textbf{R}$ otherwise.
 
-##### proof.
-_Let $M:=a_1x_1+...+a_mx_m$ and $M^c = a_{m+1}x_{m+1}+...+a_Nx_N+k$ and define the latent variable $y^L := M +  M^c$. 
-The optimal prediction rule for model $M$ predicts the state to be $\textbf{B}$ whenever 
-$P[y^L|M]\geq \frac{1}{2}$ and predicts $\textbf{R}$ otherwise._
-
-_Using the fact that for a random variable $z\sim N(\mu_z, \sigma_z)$ we have that $P[z\geq \mu_z]=\frac{1}{2}$ and 
+Using the fact that for a random variable $z\sim N(\mu_z, \sigma_z)$ we have that $P[z\geq \mu_z]=\frac{1}{2}$ and 
 noticing that $M^c$ and $y^L|M$ are Normally distributed with means $\mathbb{E}[M^c]$ and $M+\mathbb{E}[M^c]$ 
@@
 $P[y^L\geq 0 | M] < P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}$. Therefore, the optimal rule is to predict 
-$\mathbf{B}$ whenever $a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]$ and $\textbf{R}$ otherwise._
+$\mathbf{B}$ whenever $a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]$ and $\textbf{R}$ otherwise.
+$\quad \square$
+
+ \
 
@@
 
-### Polarization
+## Polarization
 We take the definition of polarization from @Haghtalab2021: two agents are said to be 
-polarized if, when observing the same realization of observables, they predict the state to be different. That is, 
-they both observe the same draw of $x=(x_1, ..., x_N)$ but one agent predicts the state to be **B** and the other 
+polarized if, when observing the realization of X, they predict the state to be different. That is, 
+they both observe the same draw of $(x_1, ..., x_N)$ but one agent predicts the state to be **B** and the other 
 predicts the state to be **R**.
@@
 Within out framework---as in @Haghtalab2021---polarization may persist even when agents have 
-access to infinite data because they might be using different models of the world. 
-Nevertheless using different models does not necessarily imply that beliefs will always be polarized, it must also be 
-the case that the realization of observables falls in a polarization zone. By *polarization zone*, we  
-mean that if the two models being used for prediction are $M_1$ and $M_2$, the realization of the 
-observables $x=(x_1, ..., x_N)$ must be such that the optimal rule for $M_1$ makes a different prediction from 
-that of $M_2$. The following 
-2-dimensional example illustrates why the use of different models is necessary but not sufficient for polarization 
-to arise.
+access to unlimited data. There are two factors that dare necessary for polarization to arise: the use of different 
+models and the realization of the observables must be such that the models will make different predictions. It is important
+to observe that even when agents are using different models, not all realiations of $X$ will produce polarized 
+predictions. The following 2-dimensional example illustrates why the use of different models is necessary but not sufficient 
+for polarization to arise.
 
-##### Example
+**Example.**
 _The optimal rule for the complex model coincides with the truth, it predicts **B** if $x_2-1\geq x_1$. 
@@
 if the value of $x_i$ is below (above) the threshold $t_i := \mu_{-i} + k$ and the predicted state is **R** otherwise.
-These prediction rules are illustrated in . As Figure \ref{example} makes clear, both rules agree on the predicted state for 
+These prediction rules are illustrated in . As Figure \ref{examp} makes clear, both rules agree on the predicted state for 
 some values of $(x_1, x_2)$ and disagree for other values. The regions in which the rules make different predictions
@@
 
-```{r example, out.width='50%', fig.align='center', fig.cap='\\label{example} Polarization regions for two single variable models'}
+```{r, out.width='50%', fig.align='center', fig.cap='\\label{examp} Polarization regions for two single variable models'}
 knitr::include_graphics(here("diagram_polariazation_2variables.png"))
@@
 to determine whether they will be polarized or not by looking at the predictions made by the optimal rule corresponding to their 
-models. Notice that the 
-polarizing observations---those that fall into a polarization zone---are different for each pair of models. Also notice that 
-our definition of the polarization zone assumes that the agents are using optimal rules. This is a very strong assumption 
-and will most likely not be the case that subjects in the experiment learn perfectly. This is a theoretical construct that allows
+models. Polarizing observations---those for which the predictions are different---are different for each pair of models. 
+Also notice that we determine whether an observation is polarizing or not by using the optimal rules. We do not expect 
+subjects in our experiment to converge to the optimal rule, this is a theoretical construct that allows
 us to formalize Prediction 2.
@@
 **Prediction 2 :** Polarization will arise more often when the theory predicts so. That is, when two agents who have 
-different models of the world, are facing a realization of observables that falls into the polarization zone corresponding 
-to their models.
+different models of the world, are facing a realization of observables that polarizing for their models.
 
@@
 
-Subjects for the experimet were recruited from the CESS lab subject pool at NYU. 
+Subjects for the experimet were recruited from NYU's CESS-lab subject pool. 
 We recruited `r data[, uniqueN(participant.code)]` undergraduate students who participated in this experiment in person 
-durinf the summer and fall of 2023. The experiment was coded using o-tree @otree and it consisted of two parts:
+during the summer and fall of 2023. The experiment was coded using oTree (@otree) and it consisted of two parts:
 Part 1 was meant to expose the subjects to the data generating process
-and allow them to learn how to predict; part 2 was meant to elicit the models of the world that subjects might have
+and allow them to learn how to predict the state; part 2 was meant to elicit the models of the world that subjects might have
 developed in part 1.
@@
 the state. We had 5 observable variables called variable 1, 2, 3, 4 and 5. The variables were independent and Normally 
-distributed with means and variances as described in TABLE.
+distributed with means and variances as described in Table \ref{params}.
 
-The state was determined by the following rule: $11x_1+6x_2+4.5x_3+2.4*x_4+k \geq 0$. The values of the coefficients were chosen
-so that all variables had a similar but unique level of informativeness when considered on their own. In this case, because
-all the variables are normally distributed, the information that they convey about the state is related to the
-variance of the modified normal $a_ix_i$. The constant $k$ was chosen so that the state was **B** in 50% of the cases. 
-The last column of TABLE shows the probability of the optimal prediction rule making a correct prediction when the model used
-considers only each variable by itself. This is the measure that we use to determine the informativeness of each variable and
-is reported in the last column of TABLE.
+The state was determined by the following linear rule: $11x_1+6x_2+4.5x_3+2.4x_4+k \geq 0$. 
+The values of the coefficients were chosen
+so that all variables had a similar but unique level of informativeness when considered on their own. The constant $k$ was 
+chosen so that the state was **B** in 50% of the cases. 
+The last column of Table \ref{params} shows the probability of the optimal prediction rule making a correct prediction 
+when for each single-variable model. This is the measure that we use to determine the informativeness of each variable.
 
@@
 
-In each round, subjects were asked to predict the state. They were given feedback on whether their prediction was correct or not.
-Both subjects within a pair were shown exactly the same realization of the variables and had to predict the same state. This allows
-us to determine whether the pair was polarized  or not for that particular observation.
+Table: Parameters for the Data Generating Process \label{params}
 
-Part 2 started immediatly after part 1 and has 70 rounds. In this part each subject was randomly allowed to use only a certain 
-number of variables in order to make their prediction. Instead of showing them the realizations of all 5 variables, they were 
-told that they can disclose up to $K$ variables, where $K$ was drawn at random from the set $\{1, 2, ..., 5\}$. K was redrawn
-independently every round and across subjects. We use the random assignment of $K$ to investigate whether the subjects are using 
+In each round, subjects were asked to predict the state. They were given feedback on whether their prediction was correct or not 
+and they had access to the entire history of the game. 
+Throughout the experiment, subjects were paired with another participant at random. Both subjects in a pair were shown exactly 
+the same realization of the variables and had to predict the same state. This allows us to determine whether the pair was 
+polarized for that particular observation. Subjects were not told that there was someone else observing the same realizations as them.
+
+Part 2 started immediatly after part 1 and had 70 rounds\footnote{In the first 2 sessions we had 40 rounds in part 2
+and the experiment was done in under 30 minutes. Because the experiment concluded in less time than expected, we increased the 
+number of rounds to 70 for the following sessions. Only 12 of our subjects participated in the sessions with 40 rounds, all others 
+were in sessions with 70 rounds for part 2}. 
+In instead of showing subjects the realizations of all 5 variables, they were 
+told that they can disclose up to $m$ variables, where $m$ was drawn at random from the set $\{1, 2, ..., 5\}$. m was redrawn
+independently across rounds and across subjects.
+
+We use the random assignment of $m$ to investigate whether the subjects are using 
 siumple or complex models. If allowing subjects to disclose an additional variable improves their prediction accuracy, 
 then it must be that they are in fact using the additional information for prediciton. If, on the other hand, they do not perform 
-better when they have access to more information, it must be that although they are disclosing an additional variable, 
-they do not correctly incorporate the infomation into their prediciton rule. Thus, if we observe that the performance of our 
-subjects does not improve when they are allowed to disclose $K+1$ variables, 
-relative to their performance when they were allowed to disclose only $K$ variables, their model of the world must be of size $K$
-or lower. This feature design allows us to test prediciton 1 and 1B and determine the number of variables that subjects are using.
+better when they have access to more information, it must be that, although they are disclosing an additional variable, 
+they do not correctly incorporate the infomation into their prediction rule. Thus, if we observe that the performance of our 
+subjects does not improve when they are allowed to disclose $m+1$ variables, 
+relative to their performance when they were allowed to disclose only $m$ variables, their model of the world must be of size $m$
+or lower. This feature design allows us to test predictions 1 and 1B and determine the number of variables that subjects are using.
 
-In order to explore the polarization prediciton, we use the fact that subjects are in fixed pairs throughout the experiment.
+In order to explore the polarization prediction, we use the fact that subjects are in fixed pairs throughout the experiment.
 Both subjects in a pair observe the same realization of all the variables in each round. Therefore, we can determine whether
 they are polarized or not by looking at whether they make the same prediction or not. We can then determine whether the theory 
-predicts that they should be polarized or not by looking at the polarization zone corresponding to the models that they are using.
-If the draw of the variables falls into the polarization zone, then the theory predicts that they should be polarized. The models
-that we use for this prediction are the ones that include the variables that they chose to reveal in that round. It could be that
-although they are revealing certain variables, they are not using them for prediction, however, we do not have a good way to 
-determine this. Instead, we use the variables that they reveal as a proxy for the variables that they are using for prediction and 
- determine the polarization zones for those models.
+predicts that they should be polarized or not by looking at what the optimal rules for those models dictate.
+We use the optimal rules that correcpond to the variables that they chose to reveal in that round. It could be that
+although they are revealing certain variables, they are not using them for prediction. Since we do not have a good method to 
+determine which variables they actually pay attention to, we use the variables that they reveal as a proxy for the variables 
+that they are using for prediction and apply the optimal rules for those models.
 
@@
 
-### Learning
+\subsection{Learning}
 In the first part of the experiment, subjects were asked to predict the state given the realization of all the variables.
@@
 the state corectly about 50% of the time, which is consistent with random guessing. The share of correct guesses increases to 
-`r 100*round(part1[subsession.round_number>15, mean(player.correct)], digits=2)`% by the end of part 1. And this last number is 
+`r 100*part1[subsession.round_number>15, mean(player.correct)]`% by the end of part 1. And this last number is 
 significantly higher than 50% 
-(p-value = `r round(t.test(part1[subsession.round_number>10, player.correct], mu=0.5)$p.value, digits=4)`). 
+(p-value = `r t.test(part1[subsession.round_number>10, player.correct], mu=0.5)$p.value`). 
 They continue to learn even after the first 20 rounds when we look at the rounds in which subjects had access to all the information.
 By the end of the experiment the rate of correct predictions is 
-`r 100*round(part2[subsession.round_number>30, mean(player.correct)], digits=2)`% which is still significantly higher than 50%
-(p-value = `r round(t.test(part2[subsession.round_number>30, player.correct], mu=0.5)$p.value, digits=4)`).
+`r 100*part2[subsession.round_number>30, mean(player.correct)]`% which is still significantly higher than 50%
+(p-value = `r t.test(part2[subsession.round_number>30, player.correct], mu=0.5)$p.value`).
 
-```{r learningAll, out.width='50%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds'}
+```{r learningAll, out.width='50%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds by performance in rounds 20 to 40'}
 knitr::include_graphics(here("computed_objects", "figures", "learning_all.png"))
 ```
+
+We also observe that there is heterogeneity in how our subjects learn. In particular, in the first 20 rounds of part 2 we are 
+able to identify two distinct groups of subjects: those who guess more than 50% of the states correctly and those who guess
+50% of fewer of them correctly. As can seen in \ref{p2groups}, these two groups have very different learning patterns 
+throughout the experiment. The group that does better than random in rounds 20 to 40 (the first 20 rounds of part 2) consists 
+of `r part2[better_random_2040==TRUE, uniqueN(participant.code)]` subjects, which account for 
+`r 100*part2[better_random_2040==TRUE, uniqueN(participant.code)]/part2[, uniqueN(participant.code)]`% of the sample. These subjects 
+improved their performance from 50% to `r 100*part2[better_random_2040==TRUE & round_number_modif<41, mean(player.correct)]`% in the 
+first 20 rounds of part 2, and continued to improve to get to 
+`r 100*part2[better_random_2040==TRUE & round_number_modif>50 & round_number_modif<61, mean(player.correct)]`% by the end of part 2. 
+Meanwhile, the group that does worse than random in rounds 20 to 40 consists of `r part2[better_random_2040==FALSE, uniqueN(participant.code)]`
+subjects and they do not seem to learn how to predict the state more accurately even after all 60 rounds of the experiment.
+Whenever it is relevant, we will show results separately for these two groups.
+
+```{r p2groups, out.width='50%', fig.align='center', fig.cap='\\label{p2groups} Share of correct predictions across rounds'}
+knitr::include_graphics(here("computed_objects", "figures", "p2_correct_rounds.png"))
+```
+
+## Simplicity
+
+In part 2, subjects were allowed to disclose up to $m$ variables before making their predictions. and $m$ was assigned at random 
+for every subject in every round. We use this random assignment to understand if allowing them to reveal an additional variable 
+improves their prediction accuracy. If it does, it must be that they are using the additional information for prediction. 
+Figure \ref{simplicity} shows the share of correct predictions by the number of variables that subjects chose to disclose.
+
+```{r simplicity, out.width='50%', fig.align='center', fig.cap='\\label{simplicity} Share of correct predictions by number of disclosed variables'}
+knitr::include_graphics(here("computed_objects", "figures", "revealed_variables_pooled.png"))
+```
+We see that there is a sharp increase when going from 1 to 2 variables, but no significant increase when going from 2 to 3 and
+beyond. This suggests that on average subjects are using mental models that consider 2 variables. However, since subjects were 
+allowed to reveal fewer variables than what was assigned to them, the estimates in \ref{simplicity} might be biased. To account 
+for this, we look at the share of correct predictions by the number of variables that subjects were allowed to disclose. The results 
+are presented in Table
+
+```{r }
+m1 <- lm(player.correct ~ C(player.number_variables) - 1, data = part2)
+m2 <- lm(player.correct ~ C(revealed_variables_count) - 1, data = part2)
+
+huxreg("Disclosed" = m2, 
+       "Allowed" = m1,
+       coefs = c("Zero variables" = "C(revealed_variables_count)0",
+                 "One variable" = "C(player.number_variables)1",
+                 "One variable" = "C(revealed_variables_count)1",
+                 "Two variables" = "C(player.number_variables)2",
+                 "Two variables" = "C(revealed_variables_count)2",
+                 "Three variables" = "C(player.number_variables)3",
+                 "Three variables" = "C(revealed_variables_count)3",
+                 "Four variables" = "C(player.number_variables)4",
+                 "Four variables" = "C(revealed_variables_count)4",
+                 "Five variables" = "C(player.number_variables)5",
+                 "Five variables" = "C(revealed_variables_count)5"), 
+        statistics = c("N. obs." = "nobs", "R squared" = "r.squared" ))
+
+```
+
+## Model Choices
+
+## Polarization Results
 

### 2023-07-13 paper/paper.tex
regressions in r

@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
+
+\setlength{\parindent}{25pt}
+\setlength{\parskip}{0pt}
+\setlength{\skip\footins}{0.25cm} % margin before footnotes
+\definecolor{myblue}{RGB}{31, 61, 122}
+\definecolor{mypink}{RGB}{204, 0, 82}
+\newtheorem{assumption}{Assumption}
+\newtheorem{proposition}{Proposition} % Use the 'proposition' counter for numbering
+
+\newcommand{\EE}[1]{\mathbb{E}\left[#1\right]}
+\newcommand{\PP}[1]{\mathbb{P}\left(#1\right)}
+\newcommand{\CP}[2]{\mathbb{P}\left(#1 \,| \, #2 \right)}
+\newcommand{\CE}[2]{\mathbb{E}\left[#1\,|\,#2 \right]}
 \usepackage[bottom]{footmisc}
@@
 \usepackage{hyperref}
+\usepackage{array}
+\usepackage{caption}
+\usepackage{graphicx}
+\usepackage{siunitx}
+\usepackage[normalem]{ulem}
+\usepackage{colortbl}
+\usepackage{multirow}
+\usepackage{hhline}
+\usepackage{calc}
+\usepackage{tabularx}
+\usepackage{threeparttable}
+\usepackage{wrapfig}
+\usepackage{adjustbox}
 \ifLuaTeX
@@
 \author{Alberto Bisin, Guillaume Frechette and Jimena Galindo}
-\date{July 11, 2023}
+\date{July 13, 2023}
 
@@
 
-TBW
-
 \hypertarget{literature-review}{%
 \section{Literature Review}\label{literature-review}}
-
-TBW
 
@@
 \hypertarget{optimal-rules}{%
-\subsubsection{Optimal Rules}\label{optimal-rules}}
+\subsection{Optimal Rules}\label{optimal-rules}}
 
@@
 
-\hypertarget{proposition.}{%
-\subparagraph{Proposition.}\label{proposition.}}
+\begin{proposition}
+Fix any model that considers $m\leq N$ variables and relabel the variable so that $x_1, ..., x_m$ are the variables 
+that the model considers and $x_{m+1}, ..., x_N$ are the ones that the model does not consider. Relabel the 
+coefficients $a = (a_1, ..., a_N)$ accordingly as well. 
+\end{proposition}
 
-\emph{Fix any model that considers \(m\leq N\) variables and relabel the
-variable so that \(x_1, ..., x_m\) are the variables that the model
-considers and \(x_{m+1}, ..., x_N\) are the ones that the model does not
-consider. Relabel the coefficients \(a = (a_1, ..., a_N)\) accordingly
-as well.}
-
-\hypertarget{proof.}{%
-\subparagraph{proof.}\label{proof.}}
-
-\emph{Let \(M:=a_1x_1+...+a_mx_m\) and
+\emph{Proof}. Let \(M:=a_1x_1+...+a_mx_m\) and
 \(M^c = a_{m+1}x_{m+1}+...+a_Nx_N+k\) and define the latent variable
@@
 the state to be \(\textbf{B}\) whenever \(P[y^L|M]\geq \frac{1}{2}\) and
-predicts \(\textbf{R}\) otherwise.}
+predicts \(\textbf{R}\) otherwise.
 
-\emph{Using the fact that for a random variable
-\(z\sim N(\mu_z, \sigma_z)\) we have that \(P[z\geq \mu_z]=\frac{1}{2}\)
-and noticing that \(M^c\) and \(y^L|M\) are Normally distributed with
-means \(\mathbb{E}[M^c]\) and \(M+\mathbb{E}[M^c]\) respectively, it is
-easy to see that
+Using the fact that for a random variable \(z\sim N(\mu_z, \sigma_z)\)
+we have that \(P[z\geq \mu_z]=\frac{1}{2}\) and noticing that \(M^c\)
+and \(y^L|M\) are Normally distributed with means \(\mathbb{E}[M^c]\)
+and \(M+\mathbb{E}[M^c]\) respectively, it is easy to see that
 \(P[y^L\geq 0 | M] \geq P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}\)
@@
 Therefore, the optimal rule is to predict \(\mathbf{B}\) whenever
-\(a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]\) and \(\textbf{R}\)
-otherwise.}
+\(a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]\) and \(\textbf{R}\) otherwise.
+\(\quad \square\)
+
+\hfill\break
 
@@
 \hypertarget{polarization}{%
-\subsubsection{Polarization}\label{polarization}}
+\subsection{Polarization}\label{polarization}}
 
 We take the definition of polarization from \citet{Haghtalab2021}: two
-agents are said to be polarized if, when observing the same realization
-of observables, they predict the state to be different. That is, they
-both observe the same draw of \(x=(x_1, ..., x_N)\) but one agent
-predicts the state to be \textbf{B} and the other predicts the state to
-be \textbf{R}.
+agents are said to be polarized if, when observing the realization of X,
+they predict the state to be different. That is, they both observe the
+same draw of \((x_1, ..., x_N)\) but one agent predicts the state to be
+\textbf{B} and the other predicts the state to be \textbf{R}.
 
 Within out framework---as in \citet{Haghtalab2021}---polarization may
-persist even when agents have access to infinite data because they might
-be using different models of the world. Nevertheless using different
-models does not necessarily imply that beliefs will always be polarized,
-it must also be the case that the realization of observables falls in a
-\emph{disagreement zone}. By disagreement zone, we simply mean that if
-the two models that are being used for prediction are \(M_1\) and
-\(M_2\), the realization of the observables \(x=(x_1, ..., x_N)\) must
-be such that \(M_1\) makes a different prediction from \(M_2\). The
-following 2-dimensional example illustrates why the use of different
-models is necessary but not sufficient for polarization to arise.
+persist even when agents have access to unlimited data. There are two
+factors that dare necessary for polarization to arise: the use of
+different models and the realization of the observables must be such
+that the models will make different predictions. It is important to
+observe that even when agents are using different models, not all
+realiations of \(X\) will produce polarized predictions. The following
+2-dimensional example illustrates why the use of different models is
+necessary but not sufficient for polarization to arise.
 
-\hypertarget{example}{%
-\subparagraph{Example}\label{example}}
-
-\emph{The optimal rule for the complex model coincides with the truth,
-it predicts \textbf{B} if \(x_2+1\geq x_1\). In addition, there are two
-simple models: one that considers only \(x_1\)---call it \(M_1\)---and
-one that considers only \(x_2\)---call it \(M_2\). The optimal rules for
-\(M_1\) and \(M_2\) are threshold rules that predict the state is
-\textbf{B} if the value of \(x_i\) is below (above) the threshold
-\(t_i := \mu_{-i} + k\) and the predicted state is \textbf{R} otherwise.
-These prediction rules are illustrated in . As Figure \ref{example}
-makes clear, both rules agree on the predicted state for some values of
-\((x_1, x_2)\) and disagree for other values. The regions in which the
-rules make different predictions are the disagreement zones.}
+\textbf{Example.} \emph{The optimal rule for the complex model coincides
+with the truth, it predicts \textbf{B} if \(x_2-1\geq x_1\). In
+addition, there are two simple models: one that considers only
+\(x_1\)---call it \(M_1\)---and one that considers only \(x_2\)---call
+it \(M_2\). The optimal rules for \(M_1\) and \(M_2\) are threshold
+rules that predict the state is \textbf{B} if the value of \(x_i\) is
+below (above) the threshold \(t_i := \mu_{-i} + k\) and the predicted
+state is \textbf{R} otherwise. These prediction rules are illustrated in
+. As Figure \ref{examp} makes clear, both rules agree on the predicted
+state for some values of \((x_1, x_2)\) and disagree for other values.
+The regions in which the rules make different predictions are the
+disagreement zones.}
 
@@
 
-\caption{\label{example} Polarization regions for two single variable models}\label{fig:example}
+\caption{\label{examp} Polarization regions for two single variable models}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 whether they will be polarized or not by looking at the predictions made
-by the optimal rule corresponding to their models. Notice that the
-polarizing observations---those that fall into a polarization zone---are
-different for each pair of models. Also notice that our definition of
-the polarization zone assumes that the agents are using optimal rules.
-This is a very strong assumption and will most likely not be the case
-that subjects in the experiment learn perfectly. This is a theoretical
-construct that allows us to formalize Prediction 2.
+by the optimal rule corresponding to their models. Polarizing
+observations---those for which the predictions are different---are
+different for each pair of models. Also notice that we determine whether
+an observation is polarizing or not by using the optimal rules. We do
+not expect subjects in our experiment to converge to the optimal rule,
+this is a theoretical construct that allows us to formalize Prediction
+2.
 
@@
 theory predicts so. That is, when two agents who have different models
-of the world, are facing a realization of observables that falls into
-the polarization zone corresponding to their models.
+of the world, are facing a realization of observables that polarizing
+for their models.
 
@@
 
-Subjects for the experimet were recruited from the CESS lab subject pool
-at NYU. We recruited 30 undergraduate students who participated in this
-experiment in person durinf the summer and fall of 2023. The experiment
-was coded using o-tree \citet{otree} and it consisted of two parts: Part
-1 was meant to expose the subjects to the data generating process and
-allow them to learn how to predict; part 2 was meant to elicit the
-models of the world that subjects might have developed in part 1.
+Subjects for the experimet were recruited from NYU's CESS-lab subject
+pool. We recruited 30 undergraduate students who participated in this
+experiment in person during the summer and fall of 2023. The experiment
+was coded using oTree (\citet{otree}) and it consisted of two parts:
+Part 1 was meant to expose the subjects to the data generating process
+and allow them to learn how to predict the state; part 2 was meant to
+elicit the models of the world that subjects might have developed in
+part 1.
 
@@
 were independent and Normally distributed with means and variances as
-described in TABLE.
+described in Table \ref{params}.
 
-The state was determined by the following rule:
-\(11x_1+6x_2+4.5x_3+2.4*x_4+k \geq 0\). The values of the coefficients
+The state was determined by the following linear rule:
+\(11x_1+6x_2+4.5x_3+2.4x_4+k \geq 0\). The values of the coefficients
 were chosen so that all variables had a similar but unique level of
-informativeness when considered on their own. In this case, because all
-the variables are normally distributed, the information that they convey
-about the state is related to the variance of the modified normal
-\(a_ix_i\). The constant \(k\) was chosen so that the state was
-\textbf{B} in 50\% of the cases. The last column of TABLE shows the
-probability of the optimal prediction rule making a correct prediction
-when the model used considers only each variable by itself. This is the
-measure that we use to determine the informativeness of each variable
-and is reported in the last column of TABLE.
+informativeness when considered on their own. The constant \(k\) was
+chosen so that the state was \textbf{B} in 50\% of the cases. The last
+column of Table \ref{params} shows the probability of the optimal
+prediction rule making a correct prediction when for each
+single-variable model. This is the measure that we use to determine the
+informativeness of each variable.
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2174}}@{}}
+\caption{Parameters for the Data Generating Process
+\label{params}}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Variable
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Mean
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Variance
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Modified Variable
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Modified Variance
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Informativeness
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
 \toprule\noalign{}
@@
 In each round, subjects were asked to predict the state. They were given
-feedback on whether their prediction was correct or not. Both subjects
-within a pair were shown exactly the same realization of the variables
-and had to predict the same state. This allows us to determine whether
-the pair was polarized or not for that particular observation.
+feedback on whether their prediction was correct or not and they had
+access to the entire history of the game. Throughout the experiment,
+subjects were paired with another participant at random. Both subjects
+in a pair were shown exactly the same realization of the variables and
+had to predict the same state. This allows us to determine whether the
+pair was polarized for that particular observation. Subjects were not
+told that there was someone else observing the same realizations as
+them.
 
-Part 2 started immediatly after part 1 and has 70 rounds. In this part
-each subject was randomly allowed to use only a certain number of
-variables in order to make their prediction. Instead of showing them the
-realizations of all 5 variables, they were told that they can disclose
-up to \(K\) variables, where \(K\) was drawn at random from the set
-\(\{1, 2, ..., 5\}\). K was redrawn independently every round and across
-subjects. We use the random assignment of \(K\) to investigate whether
-the subjects are using siumple or complex models. If allowing subjects
-to disclose an additional variable improves their prediction accuracy,
-then it must be that they are in fact using the additional information
-for prediciton. If, on the other hand, they do not perform better when
-they have access to more information, it must be that although they are
+Part 2 started immediatly after part 1 and had 70
+rounds\footnote{In the first 2 sessions we had 40 rounds in part 2
+and the experiment was done in under 30 minutes. Because the experiment concluded in less time than expected, we increased the 
+number of rounds to 70 for the following sessions. Only 12 of our subjects participated in the sessions with 40 rounds, all others 
+were in sessions with 70 rounds for part 2}. In instead of showing
+subjects the realizations of all 5 variables, they were told that they
+can disclose up to \(m\) variables, where \(m\) was drawn at random from
+the set \(\{1, 2, ..., 5\}\). m was redrawn independently across rounds
+and across subjects.
+
+We use the random assignment of \(m\) to investigate whether the
+subjects are using siumple or complex models. If allowing subjects to
+disclose an additional variable improves their prediction accuracy, then
+it must be that they are in fact using the additional information for
+prediciton. If, on the other hand, they do not perform better when they
+have access to more information, it must be that, although they are
 disclosing an additional variable, they do not correctly incorporate the
-infomation into their prediciton rule. Thus, if we observe that the
+infomation into their prediction rule. Thus, if we observe that the
 performance of our subjects does not improve when they are allowed to
-disclose \(K+1\) variables, relative to their performance when they were
-allowed to disclose only \(K\) variables, their model of the world must
-be of size \(K\) or lower. This feature design allows us to test
-prediciton 1 and 1B and determine the number of variables that subjects
+disclose \(m+1\) variables, relative to their performance when they were
+allowed to disclose only \(m\) variables, their model of the world must
+be of size \(m\) or lower. This feature design allows us to test
+predictions 1 and 1B and determine the number of variables that subjects
 are using.
 
-In order to explore the polarization prediciton, we use the fact that
+In order to explore the polarization prediction, we use the fact that
 subjects are in fixed pairs throughout the experiment. Both subjects in
@@
 whether the theory predicts that they should be polarized or not by
-looking at the polarization zone corresponding to the models that they
-are using. If the draw of the variables falls into the polarization
-zone, then the theory predicts that they should be polarized. The models
-that we use for this prediction are the ones that include the variables
-that they chose to reveal in that round. It could be that although they
-are revealing certain variables, they are not using them for prediction,
-however, we do not have a good way to determine this. Instead, we use
-the variables that they reveal as a proxy for the variables that they
-are using for prediction and determine the polarization zones for those
+looking at what the optimal rules for those models dictate. We use the
+optimal rules that correcpond to the variables that they chose to reveal
+in that round. It could be that although they are revealing certain
+variables, they are not using them for prediction. Since we do not have
+a good method to determine which variables they actually pay attention
+to, we use the variables that they reveal as a proxy for the variables
+that they are using for prediction and apply the optimal rules for those
 models.
@@
 
-\hypertarget{learning}{%
-\subsubsection{Learning}\label{learning}}
+\subsection{Learning}
 
@@
 end of part 1. And this last number is significantly higher than 50\%
-(p-value = \ensuremath{3\times 10^{-4}}). They continue to learn even
-after the first 20 rounds when we look at the rounds in which subjects
-had access to all the information. By the end of the experiment the rate
-of correct predictions is 59\% which is still significantly higher than
-50\% (p-value = 0).
+(p-value = 0.00031). They continue to learn even after the first 20
+rounds when we look at the rounds in which subjects had access to all
+the information. By the end of the experiment the rate of correct
+predictions is 59\% which is still significantly higher than 50\%
+(p-value = 0.00000019).
 
@@
 
-\caption{\label{learningAll} Share of correct predictions across rounds}\label{fig:learningAll}
+\caption{\label{learningAll} Share of correct predictions across rounds by performance in rounds 20 to 40}\label{fig:learningAll}
 \end{figure}
+
+We also observe that there is heterogeneity in how our subjects learn.
+In particular, in the first 20 rounds of part 2 we are able to identify
+two distinct groups of subjects: those who guess more than 50\% of the
+states correctly and those who guess 50\% of fewer of them correctly. As
+can seen in \ref{p2groups}, these two groups have very different
+learning patterns throughout the experiment. The group that does better
+than random in rounds 20 to 40 (the first 20 rounds of part 2) consists
+of 20 subjects, which account for 67\% of the sample. These subjects
+improved their performance from 50\% to 64\% in the first 20 rounds of
+part 2, and continued to improve to get to 69\% by the end of part 2.
+Meanwhile, the group that does worse than random in rounds 20 to 40
+consists of 10 subjects and they do not seem to learn how to predict the
+state more accurately even after all 60 rounds of the experiment.
+Whenever it is relevant, we will show results separately for these two
+groups.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/p2_correct_rounds}
+
+}
+
+\caption{\label{p2groups} Share of correct predictions across rounds}\label{fig:p2groups}
+\end{figure}
+
+\hypertarget{simplicity}{%
+\subsection{Simplicity}\label{simplicity}}
+
+In part 2, subjects were allowed to disclose up to \(m\) variables
+before making their predictions. and \(m\) was assigned at random for
+every subject in every round. We use this random assignment to
+understand if allowing them to reveal an additional variable improves
+their prediction accuracy. If it does, it must be that they are using
+the additional information for prediction. Figure \ref{simplicity} shows
+the share of correct predictions by the number of variables that
+subjects chose to disclose.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/revealed_variables_pooled}
+
+}
+
+\caption{\label{simplicity} Share of correct predictions by number of disclosed variables}\label{fig:simplicity}
+\end{figure}
+
+We see that there is a sharp increase when going from 1 to 2 variables,
+but no significant increase when going from 2 to 3 and beyond. This
+suggests that on average subjects are using mental models that consider
+2 variables. However, since subjects were allowed to reveal fewer
+variables than what was assigned to them, the estimates in
+\ref{simplicity} might be biased. To account for this, we look at the
+share of correct predictions by the number of variables that subjects
+were allowed to disclose. The results are presented in Table
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+ \label{tab:unnamed-chunk-2}
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l l}
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Disclosed \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Allowed \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Zero variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.489 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.072)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} One variable \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.534 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.535 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.023)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Two variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.602 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.598 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Three variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.584 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.600 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.029)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Four variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.599 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.611 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.025)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.028)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Five variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.576 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.590 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.030)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N. obs. \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1740\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1740\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.580\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.581\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{3}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
+\hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
+\hypertarget{model-choices}{%
+\subsection{Model Choices}\label{model-choices}}
+
+\hypertarget{polarization-results}{%
+\subsection{Polarization Results}\label{polarization-results}}
 

### 2023-07-13 analysis/part2_performance.py

@@ -19,21 +19,9 @@ part1 = pd.read_csv('data/clean/part1.csv')
 all_rounds = pd.read_csv('data/clean/all.csv')
 part2 = pd.read_csv('data/clean/part2.csv', dtype={'model': str})
 
-# count the number of guesses they got right in the first 20 rounds of part 2
-guesses2040 = part2[(part2['subsession.round_number']<=20)].groupby('participant.code').sum()['player.correct']
-# add the count column to the all_rounds table
-all_rounds = all_rounds.merge(guesses2040, on='participant.code').rename(columns={'player.correct_y':'2040_count', 'player.correct_x':'player.correct'})
-# create an indicator for whether they were better than random in the first 20 rounds of part 2
-all_rounds.loc[all_rounds['2040_count']>10, 'better_random_2040'] = 1
-all_rounds.loc[all_rounds['2040_count']<=10, 'better_random_2040'] = 0
 # plot the performance across all rounds by performance in the first 20 rounds of part 2
 sns.lmplot(data=all_rounds[all_rounds['round_number_modif']<=60], x="round_number_modif", y="player.correct", x_bins=10, hue='better_random_2040')
 plt.savefig('computed_objects/figures/p2_correct_rounds.png')
-
-# add the indicator to the part2 table
-part2 = part2.merge(guesses2040, on='participant.code').rename(columns={'player.correct_y':'2040_count', 'player.correct_x':'player.correct'})
-part2.loc[part2['2040_count']>10, 'better_random_2040'] = 1
-part2.loc[part2['2040_count']<=10, 'better_random_2040'] = 0
 
 # Accuracy wrt number of variables that they chose to reveal and by performance in part2
 fig, axs = plt.subplots()

### 2023-07-13 analysis/regressions.py

@@ -12,6 +12,7 @@ import statsmodels.formula.api as smf
 from stargazer.stargazer import Stargazer
 from linearmodels.panel import RandomEffects
 import linearmodels
+import pickle
 
 # load the data in long format for each part
 part1 = pd.read_csv('data/clean/part1.csv')
@@ -26,11 +27,7 @@ mod = smf.ols(formula='y ~ indic_1 + indic_2+ indic_3 + indic_4 + indic_5  - 1',
 
 res_assigned = mod.fit(cov_type='HC3')
 
-stargazer = Stargazer([res_assigned], )
-
-with open('computed_objects/tables/regression_assigned.txt', 'w') as f:
-    f.write(stargazer.render_latex())
-
+res_assigned.save('computed_objects/tables/regression_assigned_mod.pkl')
 
 
 # For the regression using assignment to treatment directly,

### 2023-07-13 data/scripts/clean_data.py

@@ -119,7 +119,18 @@ pairs[(pairs['p2_x1']==pairs['p1_x1'])&(pairs['p2_x2']==pairs['p1_x2'])&(pairs['
 pairs.loc[pairs['p1_guess']!=pairs['p2_guess'], 'polarized']=1
 pairs.loc[pairs['p1_guess']==pairs['p2_guess'], 'polarized']=0
 
+# count the number of guesses they got right in the first 20 rounds of part 2
+guesses2040 = part2[(part2['subsession.round_number']<=20)].groupby('participant.code').sum()['player.correct']
+# add the count column to the all_rounds table
+all_rounds = all_rounds.merge(guesses2040, on='participant.code').rename(columns={'player.correct_y':'2040_count', 'player.correct_x':'player.correct'})
+# create an indicator for whether they were better than random in the first 20 rounds of part 2
+all_rounds.loc[all_rounds['2040_count']>10, 'better_random_2040'] = 1
+all_rounds.loc[all_rounds['2040_count']<=10, 'better_random_2040'] = 0
 
+# add the indicator to the part2 table
+part2 = part2.merge(guesses2040, on='participant.code').rename(columns={'player.correct_y':'2040_count', 'player.correct_x':'player.correct'})
+part2.loc[part2['2040_count']>10, 'better_random_2040'] = 1
+part2.loc[part2['2040_count']<=10, 'better_random_2040'] = 0
 
 # Save all the data in clean format
 

### 2023-07-13 packages.py

@@ -10,9 +10,9 @@ from pathlib import Path
 import statsmodels.api as sm
 import statsmodels.formula.api as smf
 from stargazer.stargazer import Stargazer
+import pickle
 
 # load the data in long format for each part
 part1 = pd.read_csv('data/clean/part1.csv')
 all_rounds = pd.read_csv('data/clean/all.csv')
 part2 = pd.read_csv('data/clean/part2.csv', dtype={'model': str})
-

### 2023-07-13 paper/compile_paper.R

@@ -5,4 +5,4 @@ renv::restore()
 knitr::opts_chunk$set(eval=T, echo=F, warning=F, message=F, cache=F)
 
 # compile the paper
-rmarkdown::render("paper/paper.Rmd")
+rmarkdown::render("paper/paper.Rmd", output_format = "pdf_document")

### 2023-07-14 paper/paper.Rmd
regressions and hyothesis tests as scripts

@@
         number_sections: true
+        template: NULL
         keep_tex: true
@@
 ```{r dependencies, include=FALSE}
-library(data.table)
-library(here)
-library(huxtable)
-library(broom)
-renv::snapshot()
-library("languageserver")
-library(reticulate)
+source("../parameters_and_packages.R")
+```
 
+```{r setup, include=FALSE}
 # Format the numbers in the inline code output
@@
 
-```{r, out.width='50%', fig.align='center', fig.cap='\\label{examp} Polarization regions for two single variable models'}
+```{r, out.width='65%', fig.align='center', fig.cap='\\label{examp} Polarization regions for two single variable models'}
 knitr::include_graphics(here("diagram_polariazation_2variables.png"))
@@
 
-```{r learningAll, out.width='50%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds by performance in rounds 20 to 40'}
+```{r learningAll, out.width='65%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds by performance in rounds 20 to 40'}
 knitr::include_graphics(here("computed_objects", "figures", "learning_all.png"))
@@
 
-```{r p2groups, out.width='50%', fig.align='center', fig.cap='\\label{p2groups} Share of correct predictions across rounds'}
+```{r p2groups, out.width='65%', fig.align='center', fig.cap='\\label{p2groups} Share of correct predictions across rounds'}
 knitr::include_graphics(here("computed_objects", "figures", "p2_correct_rounds.png"))
@@
 Figure \ref{simplicity} shows the share of correct predictions by the number of variables that subjects chose to disclose. 
+Figure \ref{simplicity2} shows the number of variables that subjects chose to disclose by the number they we allowed to disclose.
+On average, they always disclose fewer variables than what is available to them. This is a strong indication that at least some
+of the subjects are using simple models. 
+Further evidence is provided by the accuracy of their predictions as discussed in what follows
 
-```{r simplicity, out.width='50%', fig.align='center', fig.cap='\\label{simplicity} Share of correct predictions by number of disclosed variables'}
+```{r simplicity2, out.width='65%', fig.align='center', fig.cap='\\label{simplicity2} Number of disclosed variables by number of allowed variables'}
+knitr::include_graphics(here("computed_objects", "figures", "revealed_available.png"))
+```
+
+```{r simplicity, out.width='65%', fig.align='center', fig.cap='\\label{simplicity} Share of correct predictions by number of disclosed variables'}
 knitr::include_graphics(here("computed_objects", "figures", "revealed_variables_pooled.png"))
 ```
+
 We see that there is a sharp increase when going from 1 to 2 variables, but no significant increase when going from 2 to 3 and
@@
 allowed to reveal fewer variables than what was assigned to them, the estimates in \ref{simplicity} might be biased. To account 
-for this, we look at the share of correct predictions by the number of variables that subjects were allowed to disclose. The results 
-are presented in Table
+for this, we also look at the share of correct predictions by the number of variables that subjects were allowed to disclose. The results 
+are presented in Table \ref{regression_table}. The table presents the regression estimates for the model with the assigned number of
+variables as well as the model with the number of variables that subjects chose to disclose. The last column clusters standard errors
+at the group level.
 
-```{r }
-m1 <- lm(player.correct ~ C(player.number_variables) - 1, data = part2)
-m2 <- lm(player.correct ~ C(revealed_variables_count) - 1, data = part2)
+```{r regressionTable}
+regs <- readRDS(here("computed_objects", "tables", "regressions.rds"))
 
-huxreg("Disclosed" = m2, 
-       "Allowed" = m1,
+regression_table <- huxreg("Disclosed" = regs$m2,
+                           "Allowed" = regs$m1,
+                           "Allowed" = regs$m4,
        coefs = c("Zero variables" = "C(revealed_variables_count)0",
@@
                  "Five variables" = "C(revealed_variables_count)5"),
-        statistics = c("N. obs." = "nobs", "R squared" = "r.squared" ))
+        statistics = c("N. obs." = "nobs", "R squared" = "r.squared")
+        )
+
+row <- data.frame("Clustered SE", "NO", "NO", "Group")
+regression_table <- insert_row(regression_table, row, after = 13)
+caption(regression_table) <- "Regression results, dependent variable is the share of correct predictions."
+label(regression_table) <- "regression_table"
+regression_table
+```
+
+```{r hypothesesAll}
+# load the tables
+tables <- readRDS(here("computed_objects", "tables", "hypothesis_tests.rds"))
+results_tbl1 <- tables$tbl1
+results_tbl2 <- tables$tbl2
+results_tbl4 <- tables$tbl4
+# assigned variables
+test_assigned <- results_tbl1 %>% as_huxtable() %>% set_col_width(c(0.5, 0.25, 0.25))
+label(test_assigned) <- "test_assigned"
+caption(test_assigned) <- "Hypothesis tests for the regression with the number of variables that subjects were allowed to disclose."
+
+# clustered standard errors
+test_clustered <- results_tbl4 %>% as_huxtable() %>% set_col_width(c(0.5, 0.25, 0.25))
+label(test_clustered) <- "test_clustered"
+caption(test_clustered) <- "Hypothesis tests for the regression with the number of variables that subjects were allowed to disclose and clustered standard errors."
+
+# revealed variables
+test_revealed <- results_tbl2 %>% as_huxtable() %>% set_col_width(c(0.5, 0.25, 0.25))
+label(test_revealed) <- "test_revealed"
+caption(test_revealed) <- "Hypothesis tests for the regression with the number of variables that subjects chose to disclose."
 
@@
 
+We are interested in testing whether the probability of guessing the state correctly increases when subjects are allowed to 
+disclose an additional variable. This corresponds to testing the linear hypotheses $\beta_{m+1}-\beta_m=0$ for $m=1, 2, 3, 4$.
+The results of these tests are presented in Table \ref{hypotheses}. We see that the probability of guessing the state when subjects
+are allowed to disclose 2 variables is significantly higher than when they are allowed to disclose only 1 variable. And the effect 
+of having access to the additional variable is `r results_tbl4[2, Estimate]` percentage points. Allowing 3, 4 and 5 variables has no 
+effect, which sugests that subjects are using models that consider 2 variables. All the hypothesis tests are reported in 
+Tables \ref{test_assigned},\ref{test_revealed} and \ref{test_clustered} in the appendix.
+
 ## Model Choices
+Since we now know that subjects are using models that consider 2 variables, we can look at the specific models that they are using.
+We look at two particular features: Are they choosing the most informative models? and do they improve their model choices over time?
 
@@
 
+\newpage
+\appendix
+
 # Appendix
+```{r appendix}
+test_revealed
+test_assigned
+test_clustered
+```

### 2023-07-14 paper/paper.tex
regressions and hyothesis tests as scripts

@@
 \makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \author{Alberto Bisin, Guillaume Frechette and Jimena Galindo}
-\date{July 13, 2023}
+\date{July 14, 2023}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{../diagram_polariazation_2variables}
+{\centering \includegraphics[width=0.65\linewidth]{../diagram_polariazation_2variables}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/learning_all}
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/learning_all}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/p2_correct_rounds}
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/p2_correct_rounds}
 
@@
 the share of correct predictions by the number of variables that
-subjects chose to disclose.
+subjects chose to disclose. Figure \ref{simplicity2} shows the number of
+variables that subjects chose to disclose by the number they we allowed
+to disclose. On average, they always disclose fewer variables than what
+is available to them. This is a strong indication that at least some of
+the subjects are using simple models. Further evidence is provided by
+the accuracy of their predictions as discussed in what follows
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/revealed_variables_pooled}
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/revealed_available}
+
+}
+
+\caption{\label{simplicity2} Number of disclosed variables by number of allowed variables}\label{fig:simplicity2}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/revealed_variables_pooled}
 
@@
 variables than what was assigned to them, the estimates in
-\ref{simplicity} might be biased. To account for this, we look at the
-share of correct predictions by the number of variables that subjects
-were allowed to disclose. The results are presented in Table
+\ref{simplicity} might be biased. To account for this, we also look at
+the share of correct predictions by the number of variables that
+subjects were allowed to disclose. The results are presented in Table
+\ref{regression_table}. The table presents the regression estimates for
+the model with the assigned number of variables as well as the model
+with the number of variables that subjects chose to disclose. The last
+column clusters standard errors at the group level.
 
@@
 \begin{threeparttable}
- \label{tab:unnamed-chunk-2}
+\captionsetup{justification=centering,singlelinecheck=off}
+\caption{Regression results, dependent variable is the share of correct predictions.}
+ \label{regression_table}
 \setlength{\tabcolsep}{0pt}
-\begin{tabular}{l l l}
+\begin{tabular}{l l l l}
 
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
 \arrayrulecolor{black}
@@
 \multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Disclosed \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Allowed \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Allowed \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
 \arrayrulecolor{black}
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Zero variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.489 *** \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.489 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.072)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.072)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} One variable \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.535 *** \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.534 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.535 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.534 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.023)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.033)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Two variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.598 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.602 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.602 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.023)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.028)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Two variables \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.602 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.598 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Three variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.600 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.584 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.584 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.029)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.031)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Three variables \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.584 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.600 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Four variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.611 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.599 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.599 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.029)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.028)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.025)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.042)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Four variables \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.599 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.611 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Five variables \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.590 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.576 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.576 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.025)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.028)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.030)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.040)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
-\hhline{}
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
 \arrayrulecolor{black}
 
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Five variables \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.576 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.590 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Clustered SE \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} NO\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} NO\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} Group\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.027)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.030)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
 \arrayrulecolor{black}
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N. obs. \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1740\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1740\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R squared \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.581\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.580\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.581\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.580\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
 \arrayrulecolor{black}
 
-\multicolumn{3}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
+\multicolumn{4}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
@@
 
+We are interested in testing whether the probability of guessing the
+state correctly increases when subjects are allowed to disclose an
+additional variable. This corresponds to testing the linear hypotheses
+\(\beta_{m+1}-\beta_m=0\) for \(m=1, 2, 3, 4\). The results of these
+tests are presented in Table \ref{hypotheses}. We see that the
+probability of guessing the state when subjects are allowed to disclose
+2 variables is significantly higher than when they are allowed to
+disclose only 1 variable. And the effect of having access to the
+additional variable is 0.068 percentage points. Allowing 3, 4 and 5
+variables has no effect, which sugests that subjects are using models
+that consider 2 variables. All the hypothesis tests are reported in
+Tables \ref{test_assigned},\ref{test_revealed} and \ref{test_clustered}
+in the appendix.
+
 \hypertarget{model-choices}{%
 \subsection{Model Choices}\label{model-choices}}
+
+Since we now know that subjects are using models that consider 2
+variables, we can look at the specific models that they are using. We
+look at two particular features: Are they choosing the most informative
+models? and do they improve their model choices over time?
 
@@
 
-\renewcommand\refname{Appendix}
+\newpage
+\appendix
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+\captionsetup{justification=centering,singlelinecheck=off}
+\caption{Hypothesis tests for the regression with the number of variables that subjects chose to disclose.}
+ \label{test_revealed}
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l l}
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Hypothesis \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} Estimate \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} p-value \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_0=.5 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.0106\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.883\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_1-b0=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0454\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.548\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_2-b1=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0629\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0734 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_3-b2=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.00237 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.952\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_4-b3=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0115\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.774\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_5-b4=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.0211\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.607\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+\captionsetup{justification=centering,singlelinecheck=off}
+\caption{Hypothesis tests for the regression with the number of variables that subjects were allowed to disclose.}
+ \label{test_assigned}
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l l}
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Hypothesis \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} Estimate \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} p-value \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_1=.5 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0335 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.209\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_2-b1=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0682 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0704 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_3-b2=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.0182 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.634\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_4-b3=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0159 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.67\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_5-b4=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.0231 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.53\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+\captionsetup{justification=centering,singlelinecheck=off}
+\caption{Hypothesis tests for the regression with the number of variables that subjects were allowed to disclose and clustered standard errors.}
+ \label{test_clustered}
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l l}
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Hypothesis \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} Estimate \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} p-value \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_1=.5 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0335 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.315\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_2-b1=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0682 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0273 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_3-b2=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.0182 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.6\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_4-b3=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.0159 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.686\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} b\_5-b4=0 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.0231 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.544\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
   \bibliography{bibliography.bib}

### 2023-07-14 analysis/hypothesis_tests.R

@@ -1,0 +1,100 @@
+#!/usr/bin/env Rscript
+
+# read the dependencies
+source("parameters_and_packages.R")
+
+data <- fread(here("data", "clean", "all.csv"))
+part1 <- fread(here("data", "clean", "part1.csv"))
+part2 <- fread(here("data", "clean", "part2.csv"))
+
+# read the regression models
+regs <- readRDS(here("computed_objects", "tables", "regressions.rds"))
+model1 <- regs$m1
+model2 <- regs$m2
+model4 <- lm(player.correct ~ C(player.number_variables) - 1, data = part2)
+
+# Define the individual hypotheses
+hypothesis1 <- "C(player.number_variables)1 = 0.5"
+hypothesis2 <- "C(player.number_variables)2 - C(player.number_variables)1=0"
+hypothesis3 <- "C(player.number_variables)3 - C(player.number_variables)2=0"
+hypothesis4 <- "C(player.number_variables)4 - C(player.number_variables)3=0"
+hypothesis5 <- "C(player.number_variables)5 - C(player.number_variables)4=0"
+
+model1 <- regs$m1
+test_result1 <- car::linearHypothesis(model1, hypothesis1) %>% tidy()
+test_result2 <- car::linearHypothesis(model1, hypothesis2) %>% tidy()
+test_result3 <- car::linearHypothesis(model1, hypothesis3) %>% tidy()
+test_result4 <- car::linearHypothesis(model1, hypothesis4) %>% tidy()
+test_result5 <- car::linearHypothesis(model1, hypothesis5) %>% tidy()
+
+results_tbl1 <- rbind(test_result1, test_result2, test_result3, test_result4, test_result5) %>% as.data.table()
+results_tbl1 <- results_tbl1[, .(Hypothesis = term,
+                               Estimate = estimate,
+                               `p-value` = p.value)]
+
+results_tbl1[1, 1] <- "b_1=.5"
+results_tbl1[2, 1] <- "b_2-b1=0"
+results_tbl1[3, 1] <- "b_3-b2=0"
+results_tbl1[4, 1] <- "b_4-b3=0"
+results_tbl1[5, 1] <- "b_5-b4=0"
+
+# Do it for model 4 (clustered errors)
+
+cov_matrix <- vcovCL(model4, cluster = ~group_id)
+
+# test the hypotheses with the clustered standard errors
[…]

### 2023-07-14 analysis/regressions.R

@@ -1,0 +1,21 @@
+#!/usr/bin/env Rscript
+
+# read the dependencies
+source("parameters_and_packages.R")
+
+data <- fread(here("data", "clean", "all.csv"))
+part1 <- fread(here("data", "clean", "part1.csv"))
+part2 <- fread(here("data", "clean", "part2.csv"))
+
+# regression with the number of variables that subjects were allowed to disclose
+model1 <- lm(player.correct ~ C(player.number_variables) - 1, data = part2)
+model2 <- lm(player.correct ~ C(revealed_variables_count) - 1, data = part2)
+
+# cluster errors by group
+model4 <- lm(player.correct ~ C(player.number_variables) - 1, data = part2)
+m4coeffs_cl <- coeftest(model4, vcov = vcovCL, cluster = ~group_id, save = TRUE)
+
+models <- list(m1 = model1, m2 = model2, m4 = m4coeffs_cl)
+
+# save the model
+saveRDS(models, here("computed_objects", "tables", "regressions.rds"))

### 2023-07-14 data/scripts/clean_data.py

@@ -132,6 +132,11 @@ part2 = part2.merge(guesses2040, on='participant.code').rename(columns={'player.
 part2.loc[part2['2040_count']>10, 'better_random_2040'] = 1
 part2.loc[part2['2040_count']<=10, 'better_random_2040'] = 0
 
+# add a column that indicates the group in each session
+all_rounds.loc[: , 'group_id'] = all_rounds['session.code'] + all_rounds['group.id_in_subsession'].astype(str)
+part1.loc[: , 'group_id'] = part1['session.code'] + part1['group.id_in_subsession'].astype(str)
+part2.loc[: , 'group_id'] = part2['session.code'] + part2['group.id_in_subsession'].astype(str)
+
 # Save all the data in clean format
 
 path_p1 = Path('data/clean/part1.csv')  

### 2023-07-14 parameters_and_packages.R

@@ -1,0 +1,16 @@
+# packages and parameters for the project
+library(data.table)
+library(here)
+library(huxtable)
+library(broom)
+library(magrittr)
+library(languageserver)
+library(reticulate)
+library(car) # for linearHypothesis
+library(lmtest) # for coefTest
+library(sandwich) # for clustering errors
+
+invisible(capture.output(renv::snapshot())) # keep the packages synced with the lockfile
+
+# parameters
+params = list(date_of_exp = '2023-04-01')

### 2023-07-18 paper/paper.Rmd
editing text

@@
 
-In each period, an agent observes the realization of all observable variables and predicts the state. Their flow payoff in 
+In each period, an agent observes the realization of all variables and predicts the state. Their flow payoff in 
 period $t$ is 1 if they predict the state correctly, and it is zero otherwise. Denote the prediction by $\hat{y}$.
@@
 $M_t \subseteq \{x_1, x_2, ..., x_N\}$ is the set of variables that they choose to consider in period $t$. 
-Our main object of study is the set of variables that the agent decides to consider in a given period, and the procedure by 
-which the maping or the set are updated. We refer to the set of variables that are taken into consideration, $M_t$, as the agent's 
+Our main object of study is the set of variables that the agent decides to consider in a given period. We
+abstract away from the procedure by which the maping and the set are updated. 
+We refer to the set of variables that are taken into consideration, $M_t$, as the agent's 
 *mental model*, and assume that the mapping is linear.
@@
 A mental model is *simple* if it does not include all the relevant variables, 
-*i.e.* if $M\subset \mathcal{R}$. And it is *complex* if it includes all relevant variables, $M=\mathcal{R}$. In 
-addition, mental models can be ranked in terms of their simplicity by the number of relevant variables that they consider.
+*i.e.* if $M\subset \mathcal{R}$. And it is *complex* if it includes all relevant variables, ($\mathcal{R}\subseteq M$). 
+Mental models can be ranked in terms of their simplicity by the number of relevant variables that they consider: a model that 
+considers only $x_1$ is simpler than one that considers $x_2$ and $x_3$.
 
-Whenever agents use simple models, they ignore some of the relevant information that is available to them.
-In contrast, whenever agents use complex models, they use all the relevant information available. This means that when agents 
-with complex mental models are restricted in terms of how many variables they can consider, their ability
+Agents that use simple models ignore some of the relevant information that is available to them. While agents who use comples models
+use all of the relevant information.
+This means that when agents with complex mental models are restricted in terms of how many variables they can consider, their ability
 to predict the state will be impaired. On the contrary, restricting the number of variables that an agent with a simple model can 
@@
 part2 <- fread(here("data", "clean", "part2.csv"))
-
+count_models <- fread(here("data", "clean", "count_models.csv"))
+pairs <- fread(here("data", "clean", "pairs.csv"))
 ```
@@
 disclose an additional variable. This corresponds to testing the linear hypotheses $\beta_{m+1}-\beta_m=0$ for $m=1, 2, 3, 4$.
-The results of these tests are presented in Table \ref{hypotheses}. We see that the probability of guessing the state when subjects
+We see that the probability of guessing the state when subjects
 are allowed to disclose 2 variables is significantly higher than when they are allowed to disclose only 1 variable. And the effect 
@@
 
+Conditional on the number of variables that they were allowed to disclose, subjects explore an average of 
+`r mean(count_models$number_models)` models thoughout the experiment\footnote{When there are 5 single-variable models, 10 models 
+for 2 and 3 variables each, 5 models that consider 4 variables plus the complex and the empty models for an average of 22.2 
+possible models to choose from in each round}. 
+Figure \ref{modelCount} shows the distribution of the number of models
+that subjects explored for each number of allowed variables. For the majority of subjects we do not see wide exploration across
+models, which suggests some level of stickiness in their model choices.
+
+```{r modelCount, out.width='65%', fig.align='center', fig.cap='\\label{modelCount} Distribution of the number of models explored by number of allowed variables'}
+knitr::include_graphics(here("computed_objects", "figures", "model_count.png"))
+```
+
+We also look at the informativeness of the models that subjects chose to explore. Figure \ref{modelInformativeness} shows how the 
+average informativeness of the chose models changed across rounds. For all numbers of allowed variables, the average informativeness
+stays relatively constant throughout the experiment. This suggests that subjects are not learning to choose more informative models, 
+which is also consitent with the fact that the fact that our subjecst are not exploring a wide range of models.
+
+```{r modelInformativeness, out.width='65%', fig.align='center', fig.cap='\\label{modelInformativeness} Average informativeness of the models chosen by number of allowed variables'}
+knitr::include_graphics(here("computed_objects", "figures", "p1_performance_throughout.png"))
+```
+
+If we focus our attention at only the rounds in which they could disclose up to 2 variables, we see high heterogeneity in the model 
+choices. This heterogeneity does not decrease over time. The bottom panel of Figure \ref{modelChoices} shows the distribution of 
+the models that subjects chose in such rounds. The top pannel shows the comparison between the informativeness of the model and 
+the share of correct guesses for subjects who chose that model.
+Figure \ref{modelChoicesAll} in the appendix shows the distribution for all possible models.
+
+```{r modelChoices, out.width='65%', fig.align='center', fig.cap='\\label{modelChoices} Distribution of the models chosen'}
+knitr::include_graphics(here("computed_objects", "figures", "model_choices_2vars.png"))
+```
+
+Put together with the fact that subjects continue to learn and predict better throughout the experiment, this suggests that 
+they become better at using the same mental models, and not that they are learning to use more informative models.
+
 ## Polarization Results
+As for polarization, we see that it arises in `r pairs[predicted_polarization==TRUE & round_number_modif>40, mean(polarized)]` of the guesses in which 
+it was predicted. While it arises only in `r pairs[predicted_polarization==FALSE & round_number_modif>40, mean(polarized)]` of the guesses in which it 
+was not predicted\footnote{We use all the data after 20 rounds of part 2 to allow for some learning on how to select models}. 
+This difference is statistically significant 
+(p-value = `r t.test(pairs[predicted_polarization==TRUE & round_number_modif>40, polarized], pairs[predicted_polarization==FALSE & round_number_modif>40, polarized])$p.value`).
+
+```{r polarization, out.width='65%', fig.align='center', fig.cap='\\label{polarization} Polarization by whether it was predicted or not'}
+knitr::include_graphics(here("computed_objects", "figures", "polarization.png"))
+```
+
+The prediction is that 100% of the pairs would be polarized whenever the optimal rules for their models predict polarization and 
+none of them would be when it is not predicted. There are two factors that may influence the difference with our results. The first 
+is that subjects might not have converged to the optimal rules for the models that they selected. This would induce discrepancy with 
+the prediciton. The other is that we use take the varaibles that they chose to disclose in that round as the model that they are 
+using. However, from the discussion in the previous section we know that even though subjects are disclosing several variables, 
+they do not necessarily use the extra information. This would induce discrepancy with the prediction as well. Nonetheless, we 
+take the fact that polarization is higher when predicted by the theory as an indication that subjects are using simple models to 
+some extent. Figure \ref{polarizationRounds} shows the share of polarized pairs by whether polarization was predicted or not 
+across rounds.
+
+```{r polarizationRounds, out.width='65%', fig.align='center', fig.cap='\\label{polarizationRounds} Polarization by whether it was predicted or not'}
+knitr::include_graphics(here("computed_objects", "figures", "polarization_rounds_predicted.png"))
+```
 
@@
 # Appendix
-```{r appendix}
+```{r TestTables}
 test_revealed
@@
 ```
+
+```{r modelChoicesAll, out.width='65%', fig.align='center', fig.cap='\\label{modelChoicesAll} Distribution of the models chosen'}
+knitr::include_graphics(here("computed_objects", "figures", "model_choices_all.png"))
+```

### 2023-07-18 paper/paper.tex
editing text

@@
 \author{Alberto Bisin, Guillaume Frechette and Jimena Galindo}
-\date{July 14, 2023}
+\date{July 16, 2023}
 
@@
 additional variable. This corresponds to testing the linear hypotheses
-\(\beta_{m+1}-\beta_m=0\) for \(m=1, 2, 3, 4\). The results of these
-tests are presented in Table \ref{hypotheses}. We see that the
+\(\beta_{m+1}-\beta_m=0\) for \(m=1, 2, 3, 4\). We see that the
 probability of guessing the state when subjects are allowed to disclose
@@
 
+Conditional on the number of variables that they were allowed to
+disclose, subjects explore an average of 2.8 models thoughout the
+experiment\footnote{When there are 5 single-variable models, 10 models 
+for 2 and 3 variables each, 5 models that consider 4 variables plus the complex and the empty models for an average of 22.2 
+possible models to choose from in each round}. Figure \ref{modelCount}
+shows the distribution of the number of models that subjects explored
+for each number of allowed variables. For the majority of subjects we do
+not see wide exploration across models, which suggests some level of
+stickiness in their model choices.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/model_count}
+
+}
+
+\caption{\label{modelCount} Distribution of the number of models explored by number of allowed variables}\label{fig:modelCount}
+\end{figure}
+
+We also look at the informativeness of the models that subjects chose to
+explore. Figure \ref{modelInformativeness} shows how the average
+informativeness of the chose models changed across rounds. For all
+numbers of allowed variables, the average informativeness stays
+relatively constant throughout the experiment. This suggests that
+subjects are not learning to choose more informative models, which is
+also consitent with the fact that the fact that our subjecst are not
+exploring a wide range of models.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/p1_performance_throughout}
+
+}
+
+\caption{\label{modelInformativeness} Average informativeness of the models chosen by number of allowed variables}\label{fig:modelInformativeness}
+\end{figure}
+
+If we focus our attention at only the rounds in which they could
+disclose up to 2 variables, we see high heterogeneity in the model
+choices. This heterogeneity does not decrease over time. The bottom
+panel of Figure \ref{modelChoices} shows the distribution of the models
+that subjects chose in such rounds. The top pannel shows the comparison
+between the informativeness of the model and the share of correct
+guesses for subjects who chose that model. Figure \ref{modelChoicesAll}
+in the appendix shows the distribution for all possible models.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/model_choices_2vars}
+
+}
+
+\caption{\label{modelChoices} Distribution of the models chosen}\label{fig:modelChoices}
+\end{figure}
+
+Put together with the fact that subjects continue to learn and predict
+better throughout the experiment, this suggests that they become better
+at using the same mental models, and not that they are learning to use
+more informative models.
+
 \hypertarget{polarization-results}{%
 \subsection{Polarization Results}\label{polarization-results}}
+
+As for polarization, we see that it arises in 0.5 of the guesses in
+which it was predicted. While it arises only in 0.41 of the guesses in
+which it was not
+predicted\footnote{We use all the data after 20 rounds of part 2 to allow for some learning on how to select models}.
+This difference is statistically significant (p-value = 0.03).
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/polarization}
+
+}
+
+\caption{\label{polarization} Polarization by whether it was predicted or not}\label{fig:polarization}
+\end{figure}
+
+The prediction is that 100\% of the pairs would be polarized whenever
+the optimal rules for their models predict polarization and none of them
+would be when it is not predicted. There are two factors that may
+influence the difference with our results. The first is that subjects
+might not have converged to the optimal rules for the models that they
+selected. This would induce discrepancy with the prediciton. The other
+is that we use take the varaibles that they chose to disclose in that
+round as the model that they are using. However, from the discussion in
+the previous section we know that even though subjects are disclosing
+several variables, they do not necessarily use the extra information.
+This would induce discrepancy with the prediction as well. Nonetheless,
+we take the fact that polarization is higher when predicted by the
+theory as an indication that subjects are using simple models to some
+extent. Figure \ref{polarizationRounds} shows the share of polarized
+pairs by whether polarization was predicted or not across rounds.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/polarization_rounds_predicted}
+
+}
+
+\caption{\label{polarizationRounds} Polarization by whether it was predicted or not}\label{fig:polarizationRounds}
+\end{figure}
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=0.65\linewidth]{../computed_objects/figures/model_choices_all}
+
+}
+
+\caption{\label{modelChoicesAll} Distribution of the models chosen}\label{fig:modelChoicesAll}
+\end{figure}
+
   \bibliography{bibliography.bib}

### 2023-07-18 analysis/model_exploration.py

@@ -37,4 +37,8 @@ plt.title('Number of models that were used at least once by a subject')
 plt.savefig('computed_objects/figures/model_count.png') 
 
 
+path_models = Path('data/clean/count_models.csv')  
+path_models.parent.mkdir(parents=True, exist_ok=True)  
+count_models.to_csv(path_models)
+
 

### 2023-07-18 analysis/polarization.py

@@ -20,103 +20,6 @@ all_rounds = pd.read_csv('data/clean/all.csv')
 part2 = pd.read_csv('data/clean/part2.csv', dtype={'model': str})
 pairs = pd.read_csv('data/clean/pairs.csv')
 
-# set the parameters of the function that determines the state of the world
-#coefficients
-a1 = 11
-a2 = 6
-a3 = 4.5
-a4 = 2.4
-k = 50.5
-
-# means
-m1 = 0
-m2 = -10
-m3 = 5
-m4 = -5
-
-# define a function that predicts the state of the world conditional on the model chosen by the subject. 
-# The model chosen is the revealed variables (1) and the non-revealed variables (0) concatenated in a string.
-
-def predict(model, x1, x2, x3, x4):
-    # model is a string with 5 characters that indicates if each variable is revealed (1) or not (0)
-    # x1, x2, x3, x4 are the realized values of the variables
-    if   model == '10000' or model == '10001':
-        y_latent = a1*x1+a2*m2+a3*m3+a4*m4+k
-        
-    elif model == '01000' or model == '01001':
-        y_latent = a1*m1+a2*x2+a3*m3+a4*m4+k
-        
-    elif model == '00100' or model == '00101':
-        y_latent = a1*m1+a2*m2+a3*x3+a4*m4+k
-    
-    elif model == '00010' or model == '00011':
-        y_latent = a1*m1+a2*m2+a3*m3+a4*x4+k
-        
-    elif model == '11000' or model == '11001':
-        y_latent = a1*x1+a2*x2+a3*m3+a4*m4+k
-        
-    elif model == '10100' or model == '10101':
-        y_latent = a1*x1+a2*m2+a3*x3+a4*m4+k
-    
-    elif model == '10010' or model == '10011':
-        y_latent = a1*x1+a2*m2+a3*m3+a4*x4+k
-    
-    elif model == '01100' or model == '01101':
-        y_latent = a1*m1+a2*x2+a3*x3+a4*m4+k
-        
-    elif model == '01010' or model == '01011':
-        y_latent = a1*m1+a2*x2+a3*m3+a4*x4+k
-        
-    elif model == '00110' or model == '00111':
-        y_latent = a1*m1+a2*m2+a3*m3+a4*x4+k
-        
-    elif model == '11100' or model == '11101':
-        y_latent = a1*x1+a2*x2+a3*x3+a4*m4+k
-        
[…]

### 2023-07-18 data/scripts/clean_data.py

@@ -6,6 +6,7 @@ import pandas as pd
 import numpy as np
 import scipy as sp
 from pathlib import Path
+import random
 
 # load the data in long format for each part
 part1 = pd.read_csv('data/raw/Part1_2023-06-05.csv')
@@ -136,6 +137,108 @@ part2.loc[part2['2040_count']<=10, 'better_random_2040'] = 0
 all_rounds.loc[: , 'group_id'] = all_rounds['session.code'] + all_rounds['group.id_in_subsession'].astype(str)
 part1.loc[: , 'group_id'] = part1['session.code'] + part1['group.id_in_subsession'].astype(str)
 part2.loc[: , 'group_id'] = part2['session.code'] + part2['group.id_in_subsession'].astype(str)
+
+######
+# predict polarization
+######
+
+# set the parameters of the function that determines the state of the world
+#coefficients
+a1 = 11
+a2 = 6
+a3 = 4.5
+a4 = 2.4
+k = 50.5
+
+# means
+m1 = 0
+m2 = -10
+m3 = 5
+m4 = -5
+
+# define a function that predicts the state of the world conditional on the model chosen by the subject. 
+# The model chosen is the revealed variables (1) and the non-revealed variables (0) concatenated in a string.
+
+def predict(model, x1, x2, x3, x4):
+    # model is a string with 5 characters that indicates if each variable is revealed (1) or not (0)
+    # x1, x2, x3, x4 are the realized values of the variables
+    if   model == '10000' or model == '10001':
+        y_latent = a1*x1+a2*m2+a3*m3+a4*m4+k
+        
+    elif model == '01000' or model == '01001':
+        y_latent = a1*m1+a2*x2+a3*m3+a4*m4+k
+        
+    elif model == '00100' or model == '00101':
+        y_latent = a1*m1+a2*m2+a3*x3+a4*m4+k
+    
+    elif model == '00010' or model == '00011':
+        y_latent = a1*m1+a2*m2+a3*m3+a4*x4+k
+        
+    elif model == '11000' or model == '11001':
+        y_latent = a1*x1+a2*x2+a3*m3+a4*m4+k
+        
+    elif model == '10100' or model == '10101':
+        y_latent = a1*x1+a2*m2+a3*x3+a4*m4+k
+    
+    elif model == '10010' or model == '10011':
+        y_latent = a1*x1+a2*m2+a3*m3+a4*x4+k
+    
[…]

### 2023-07-24 paper/paper.Rmd
definition of optimal edited

@@
 ## Optimal Rules
-Conditional of using a particular mental model, the most accurate prediction rule that the agent can use is one that 
-predicts the state to be **B** whenever, given the observables under consideration, **B** is more likely. 
-And it predicts **R** whenever **R** is more likely. Because we abstract away from the procedure through which the rule is updated, 
-we will make predictions under the assumption that the agent has learned the parameters perfectly. That is, 
+Regardless of the particular model used by an agent agent, hty can maximize the probability of guessing the state correctly by 
+predicting the state to be **B** whenever it is more likely to be the true state than **R**. 
+In computing the relevant probabilities, uses only the variables that their mental model considers and hence is sometimes wrong.
+
+Because we abstract away from the procedure through which the rule is updated, 
+we will make predictions under the assumption that the agent has learned the parameters of their mental model perfectly. That is, 
 they have seen enough information that, under an updating procedure that converges to the truth---for example, 
 correctly specified regression model or Bayes rule---they would have converged to the true parameters already. This provides an 
-upper bound for how well any model can perform.
+upper bound for how well thier model can perform.In addition, we assume that the mental model is a linear rule that partitions 
+the space under consideration into two.
 
-We refer to a prediction rule as an *Optimal Rule* if it is the rule that a fully rational agent would 
-converge to with perfect Bayesian learning. In this context, Bayesian learning corresponds to using Bayes rule to 
-learn the coefficients $a = (a_1, a_2, ..., a_N)$ as well as the constant $K$. This means that in an $N$ dimensional 
-space, the agent must learn $N+1$ parameters which corresponds to learning the halfspace in which the state 
-is **B**. The optimal rule for the complex model coincides with the true data-generating process when learning 
-is perfect and the parameters are identified.
-
-An optimal rule conditional on model $M$ is the rule that, when restricted to the 
-variables being considered, does the best at predicting the state. In this case, the agent is learning only 
-the parameters that pertain the variables in $M$ plus a constant term. The following proposition illustrates exactly 
-what the optimal prediction rules look like when the agent is able to learn perfectly.
+For any given model, the *Optimal Rule* maps the varaibles under consideration to the most likely state. 
+The optimal rule for the complex modelcoincides with the true data-generating process and predicts the state correctly 100%
+of the time, while the optimal rules for simple models maximize the chance of guessing correctly conditonal on limited information.
 
